### 2016-07-27 Table1.tex
adding figure 1 and table 1

\begin{table}
\rowcolors{2}{white}{gray!25}
\begin{center}
\caption{Programs for identifying genomic regions under introgression} \label{tab:tools}
\begin{tabular}{llll}
\\\toprule  
\rowcolor{white}
{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule

\rowcolor{gray!25}
{\emph{\bf chromosome paiting}} &   &   &   \\
\rowcolor{gray!25}
Hapmix	& phased haplotype; reference panel		& Price et al. 2009	&  Hufford et al. 2013;  Suarez et al. 2016 \\ 
\rowcolor{gray!25}
RASPberry &	phased haplotype &	Wegmann et al. 2011	 & Christe et al. 2016 \\
\rowcolor{gray!25}
MultiMix & phased/unphased genotype; reference panel &	Churchhouse and Marchini 2013 &	Eyheramendy et al. 2015 \\
\rowcolor{gray!25}
PCAdmix	 & phased haplotype	 & Brisbin et al. 2012	 & Moreno et al. 2014; Pugach et al. 2016 \\
\rowcolor{gray!25}
LAMP  &	phased haplotypes; reference panel	 & Sankararaman et al. 2008	 & Patterson et al. 2012 \\

\rowcolor{white}
{\emph{\bf phylogenetic relationship}} &   &   &   \\
\rowcolor{white}
ABBA-BABA/D-statistics	 & biallelic SNP  &	Durand et al. 2011	 &  Heliconius Genome Consortium 2012 \\
\rowcolor{white}
fd statistic &	biallelic SNP &	Martin et al. 2015  &	Malinsky et al. 2015; Zhang et al. 2016 \\ 
\rowcolor{white}
five taxon D statistics	& biallelic SNP	&  Pease and Hahn 2015	& Fontaine et al. 2015; Pease et al. 2016 \\

\rowcolor{gray!25}
{\emph{\bf divergence}} &   &   &   \\
\rowcolor{gray!25}
Gmin &	biallelic SNP	&  Geneva et al. 2015	&  Kingan et al. 2015 \\
\rowcolor{gray!25}
RNDmin	& phased haplotype	& Rosenzweig et al. 2016	&  NA \\

\rowcolor{white}
{\emph{\bf Bayesian approaches}} &   &   &   \\
\rowcolor{white}
\rowcolor{white}

\end{tabular}
\end{center}
\end{table}