### 2017-07-05 bib_gj.tex

@article{Ross-Ibarra2007,
author = {Ross-Ibarra, Jeffrey and Morrell, Peter L. and Gaut, Brandon S.}, 
title = {Plant domestication, a unique opportunity to identify the genetic basis of adaptation},
volume = {104}, 
number = {suppl 1}, 
pages = {8641-8648}, 
year = {2007}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@article{Zeder2006,
	Author = {Zeder, Melinda A. and Emshwiller, Eve and Smith, Bruce D. and Bradley, Daniel G.},
	Journal = {Trends in Genetics},
	Number = {3},
	Pages = {139--155},
	Title = {Documenting domestication: the intersection of genetics and archaeology},
	Volume = {22},
	Year = {2006}}

@article{Gavrilenko2013,
	Author = {Gavrilenko, Tatjana and Antonova, Olga and Shuvalova, Anna and Krylova, Ekaterina and Alpatyeva, Natalia and Spooner, David M. and Novikova, Lubov},
	Journal = {Genetic Resources and Crop Evolution},
	Number = {7},
	Pages = {1997--2015},
	Title = {Genetic diversity and origin of cultivated potatoes based on plastid microsatellite polymorphism},
	Volume = {60},
	Year = {2013}
}

@article{Dempewolf2017,
	Author = {Dempewolf, Hannes and Baute, Gregory and Anderson, Justin and Kilian, Benjamin and Smith, Chelsea and Guarino, Luigi},
	Pages = {1070--1082},
	Title = {Past and Future Use of Wild Relatives in Crop Breeding},
	Volume = {57},
	Year = {2017}
}

@article{wang2017,
  title={The interplay of demography and selection during maize domestication and expansion},
  author={Wang, Li and Beissinger, Timothy Mathes and Lorant, Anne and Ross-Ibarra, Claudia and Ross-Ibarra, Jeffrey and Hufford, Matthew},
  journal={bioRxiv},
  pages={114579},
  year={2017},
  publisher={Cold Spring Harbor Labs Journals}
}

@book{Harlan1992,
	Address = {Madison, WI},
	Author = {Harlan, Jack R. },
	Date-Added = {2017-07-03 15:15:48 +0000},
	Date-Modified = {2017-07-03 15:15:48 +0000},
	Doi = {10.2135/1992.cropsandman},
	Isbn = {978-0-89118-565-9},
	La = {English},
	Publisher = {American Society of Agronomy, Crop Science Society of America},
	Title = {Crops \& Man},
	Ty = {BOOK},
	Url = {http://dx.doi.org/10.2135/1992.cropsandman},
	Year = {1992},
	Bdsk-Url-1 = {http://dx.doi.org/10.2135/1992.cropsandman}
}

@article{li2011,
  title={Inference of human population history from individual whole-genome sequences},
  author={Li, Heng and Durbin, Richard},
  journal={Nature},
  volume={475},
  number={7357},
  pages={493--496},
  year={2011},
  publisher={Nature Publishing Group}
}

@article{zhou2017,
  title={The Evolutionary Genomics of Grape (Vitis vinifera ssp. vinifera) Domestication},
  author={Zhou, Yongfeng and Massonnet, Melanie and Sanjak, Jaleal and Cantu, Dario and Gaut, Brandon S},
  journal={bioRxiv},
  pages={146373},
  year={2017},
  publisher={Cold Spring Harbor Labs Journals}
}

@article{Fuller2014,
author = {Fuller, Dorian Q and Denham, Tim and Arroyo-Kalin, Manuel and Lucas, Leilani and Stevens, Chris J. and Qin, Ling and Allaby, Robin G. and Purugganan, Michael D.}, 
title = {Convergent evolution and parallelism in plant domestication revealed by an expanding archaeological record},
volume = {111}, 
number = {17}, 
pages = {6147-6152}, 
year = {2014}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@article {PBI:PBI207,
author = {Messeguer, Joaquima and Peñas, Gisela and Ballester, Jordi and Bas, Marta and Serra, Joan and Salvia, Jordi and Palaudelmàs, Montserrat and Melé, Enric},
title = {Pollen-mediated gene flow in maize in real situations of coexistence},
journal = {Plant Biotechnology Journal},
volume = {4},
number = {6},
publisher = {Blackwell Publishing Ltd},
issn = {1467-7652},
url = {http://dx.doi.org/10.1111/j.1467-7652.2006.00207.x},
doi = {10.1111/j.1467-7652.2006.00207.x},
pages = {633--645},
keywords = {coexistence, maize, pollen gene flow},
year = {2006},
}

@article{wang2017asian,
  title={Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice},
  author={Wang, Hongru and Vieira, Filipe G and Crawford, Jacob E and Chu, Chengcai and Nielsen, Rasmus},
  journal={Genome Research},
  volume={27},
  number={6},
  pages={1029--1038},
  year={2017},
  publisher={Cold Spring Harbor Lab}
}

@article{morrell2007genetic,
  title={Genetic evidence for a second domestication of barley (Hordeum vulgare) east of the Fertile Crescent},
  author={Morrell, Peter L and Clegg, Michael T},
  journal={Proceedings of the national academy of sciences},
  volume={104},
  number={9},
  pages={3289--3294},
  year={2007},
  publisher={National Acad Sciences}
}

@article{dai2012tibet,
  title={Tibet is one of the centers of domestication of cultivated barley},
  author={Dai, Fei and Nevo, Eviatar and Wu, Dezhi and Comadran, Jordi and Zhou, Meixue and Qiu, Long and Chen, Zhonghua and Beiles, Avigdor and Chen, Guoxiong and Zhang, Guoping},
  journal={Proceedings of the National Academy of Sciences},
  volume={109},
  number={42},
  pages={16969--16973},
  year={2012},
  publisher={National Acad Sciences}
}

@article{vaughan2008evolving,
  title={The evolving story of rice evolution},
  author={Vaughan, Duncan A and Lu, Bao-Rong and Tomooka, Norihiko},
  journal={Plant science},
  volume={174},
  number={4},
  pages={394--408},
  year={2008},
  publisher={Elsevier}
}

@article{broich1981evolutionary,
  title={Evolutionary studies of the soybean: the frequency and distribution of alleles among collections of Glycine max and G. soja of various origin},
  author={Broich, Steven L and Palmer, Reid G},
  journal={Euphytica},
  volume={30},
  number={1},
  pages={55--64},
  year={1981},
  publisher={Springer}
}

@article{wendel1992genetic,
  title={Genetic diversity in Gossypium hirsutum and the origin of upland cotton},
  author={Wendel, Jonathan F and Brubaker, Curt L and Percival, A Edward},
  journal={American Journal of Botany},
  pages={1291--1310},
  year={1992},
  publisher={JSTOR}
}

@article{leotard2009phylogeography,
  title={Phylogeography and the origin of cassava: new insights from the northern rim of the Amazonian basin},
  author={L{\'e}otard, Guillaume and Duputi{\'e}, Anne and Kjellberg, Finn and Douzery, Emmanuel JP and Debain, Chantal and De Granville, Jean-Jacques and McKey, Doyle},
  journal={Molecular Phylogenetics and Evolution},
  volume={53},
  number={1},
  pages={329--334},
  year={2009},
  publisher={Academic Press}
}

@article{gebrekidan1982utilization,
  title={Utilization of germplasm in sorghum improvement},
  author={Gebrekidan, Brhane},
  journal={Sorghum in the eighties. Patancheru, India: International Crops Research Institute for the Semi-Arid Tropics},
  pages={335--345},
  year={1982}
}

@article{robertson2007genetic,
  title={Genetic resources of tomato (Lycopersicon esculentum Mill.) and wild relatives},
  author={Robertson, LARRY D and Labate, JOANNE A and Razdan, MK and Mattoo, AK},
  journal={Genetic Improvement of Solanaceous Crops. Tomato},
  volume={2},
  pages={25--75},
  year={2007}
}

@article{rick1958role,
  title={The role of natural hybridization in the derivation of cultivated tomatoes of western South America},
  author={Rick, Charles M},
  journal={Economic Botany},
  volume={12},
  number={4},
  pages={346--367},
  year={1958},
  publisher={Springer}
}

@article{fujita1997extent,
  title={The extent of natural cross-pollination in wild soybean (Glycine soja)},
  author={Fujita, R and Ohara, M and Okazaki, K and Shimamoto, Y},
  journal={Journal of Heredity},
  volume={88},
  number={2},
  pages={124--128},
  year={1997},
  publisher={Am Genetic Assoc}
}

@article{concibido2003introgression,
  title={Introgression of a quantitative trait locus for yield from Glycine soja into commercial soybean cultivars},
  author={Concibido, V and La Vallee, B and Mclaird, P and Pineda, N and Meyer, J and Hummel, L and Yang, J and Wu, K and Delannay, X},
  journal={Theoretical and Applied Genetics},
  volume={106},
  number={4},
  pages={575--582},
  year={2003},
  publisher={Springer}
}

@article{Meyer2016,
	Author = {Meyer, Rachel S and Choi, Jae Young and Sanches, Michelle and Plessis, Anne and Flowers, Jonathan M and Amas, Junrey and Dorph, Katherine and Barretto, Annie and Gross, Briana and Fuller, Dorian Q and Bimpong, Isaac Kofi and Ndjiondjop, Marie-Noelle and Hazzouri, Khaled M and Gregorio, Glenn B and Purugganan, Michael D},
	Journal = {Nat Genet},
	Month = {09},
	Number = {9},
	Pages = {1083--1088},
	Title = {Domestication history and geographical adaptation inferred from a SNP map of African rice},
	Volume = {48},
	Year = {2016}}

@inproceedings{besnard2013complex,
  title={The complex history of the olive tree: from Late Quaternary diversification of Mediterranean lineages to primary domestication in the northern Levant},
  author={Besnard, G and Khadari, B and Navascu{\'e}s, M and Fern{\'a}ndez-Mazuecos, M and El Bakkali, Ahmed and Arrigo, N and Baali-Cherif, D and de Caraffa, V Brunini-Bronzini and Santoni, S and Vargas, P and others},
  booktitle={Proc. R. Soc. B},
  volume={280},
  number={1756},
  pages={20122833},
  year={2013},
  organization={The Royal Society}
}

@inproceedings{besnard2013complex,
  title={The complex history of the olive tree: from Late Quaternary diversification of Mediterranean lineages to primary domestication in the northern Levant},
  author={Besnard, G and Khadari, B and Navascu{\'e}s, M and Fern{\'a}ndez-Mazuecos, M and El Bakkali, Ahmed and Arrigo, N and Baali-Cherif, D and de Caraffa, V Brunini-Bronzini and Santoni, S and Vargas, P and others},
  booktitle={Proc. R. Soc. B},
  volume={280},
  number={1756},
  pages={20122833},
  year={2013},
  organization={The Royal Society}
}

@article{percy1990allozyme,
  title={Allozyme evidence for the origin and diversification of Gossypium barbadense L.},
  author={Percy, RG and Wendel, Jonathan F},
  journal={Theoretical and Applied Genetics},
  volume={79},
  number={4},
  pages={529--542},
  year={1990},
  publisher={Springer}
}

@article{wendel1990allozyme,
  title={Allozyme diversity and introgression in the Galapagos Islands endemic Gossypium darwinii and its relationship to continental G. barbadense},
  author={Wendel, Jonathan F and Percy, Richard G},
  journal={Biochemical Systematics and Ecology},
  volume={18},
  number={7},
  pages={517--528},
  year={1990},
  publisher={Elsevier}
}

@article{fryxell1979natural,
  title={The natural history ofthe cotton tribe},
  author={Fryxell, PA},
  journal={Texas A},
  year={1979}
}

@article{wendel1989genetic,
  title={Genetic diversity, introgression, and independent domestication of old world cultivated cottons},
  author={Wendel, Jonathan F and Olson, Paul D and McD, James},
  journal={American Journal of Botany},
  pages={1795--1806},
  year={1989},
  publisher={JSTOR}
}

@article{abdalla2001genetic,
  title={Genetic diversity and relationships of diploid and tetraploid cottons revealed using AFLP},
  author={Abdalla, AM and Reddy, OUK and El-Zik, KM and Pepper, AE},
  journal={Theoretical and Applied Genetics},
  volume={102},
  number={2-3},
  pages={222--229},
  year={2001},
  publisher={Springer}
}

@article{gunn2011independent,
  title={Independent origins of cultivated coconut (Cocos nucifera L.) in the old world tropics},
  author={Gunn, Bee F and Baudouin, Luc and Olsen, Kenneth M},
  journal={Plos one},
  volume={6},
  number={6},
  pages={e21143},
  year={2011},
  publisher={Public Library of Science}
}

@article{broich1981evolutionary,
  title={Evolutionary studies of the soybean: the frequency and distribution of alleles among collections of Glycine max and G. soja of various origin},
  author={Broich, Steven L and Palmer, Reid G},
  journal={Euphytica},
  volume={30},
  number={1},
  pages={55--64},
  year={1981},
  publisher={Springer}
}

@article{kwon1972studies,
  title={Studies on the diversity of seed weight in the Korean soybean land races and wild soybean},
  author={Kwon, SH and Im, KH and Kim, JR},
  journal={Korean J Breed},
  volume={4},
  number={1},
  pages={70--74},
  year={1972}
}

@article{singh1988genomic,
  title={The genomic relationship between Glycine max (L.) Merr. and G. soja Sieb. and Zucc. as revealed by pachytene chromosome analysis},
  author={Singh, RJ and Hymowitz, T},
  journal={Theoretical and Applied Genetics},
  volume={76},
  number={5},
  pages={705--711},
  year={1988},
  publisher={Springer}
}

@article{ray2003soybean,
  title={Soybean natural cross-pollination rates under field conditions},
  author={Ray, Jeffery D and Kilen, Thomas C and Abel, Craig A and Paris, Robert L},
  journal={Environmental Biosafety Research},
  volume={2},
  number={2},
  pages={133--138},
  year={2003},
  publisher={EDP Sciences}
}

@article{ellstrand1999gene,
  title={Gene flow and introgression from domesticated plants into their wild relatives},
  author={Ellstrand, Norman C and Prentice, Honor C and Hancock, James F},
  journal={Annual review of Ecology and Systematics},
  pages={539--563},
  year={1999},
  publisher={JSTOR}
}

@article{quist2001transgenic,
  title={Transgenic DNA introgressed into traditional maize landraces in Oaxaca, Mexico},
  author={Quist, David and Chapela, Ignacio H},
  journal={Nature},
  volume={414},
  number={6863},
  pages={541--543},
  year={2001},
  publisher={Nature Publishing Group}
}

@article{doebley1990molecular,
  title={Molecular evidence for gene flow among Zea species},
  author={Doebley, John},
  journal={BioScience},
  pages={443--448},
  year={1990},
  publisher={JSTOR}
}

@article{viaud2008spatial,
  title={Spatial sensitivity of maize gene-flow to landscape pattern: a simulation approach},
  author={Viaud, Val{\'e}rie and Monod, Herv{\'e} and Lavigne, Claire and Angevin, Fr{\'e}d{\'e}rique and Adamczyk, Katarzyna},
  journal={Landscape Ecology},
  volume={23},
  number={9},
  pages={1067--1079},
  year={2008},
  publisher={Springer}
}

@article{baltazar2005pollination,
  title={Pollination between maize and teosinte: an important determinant of gene flow in Mexico},
  author={Baltazar, Baltazar M and de Jesus Sanchez-Gonzalez, Jose and de la Cruz-Larios, Lino and Schoper, John B},
  journal={Theoretical and Applied Genetics},
  volume={110},
  number={3},
  pages={519--526},
  year={2005},
  publisher={Springer}
}

@article{galinat1973intergenomic,
  title={Intergenomic mapping of maize, teosinte and Tripsacum},
  author={Galinat, Walton C},
  journal={Evolution},
  pages={644--655},
  year={1973},
  publisher={JSTOR}
}

@article{ellstrand2007spontaneous,
  title={Spontaneous hybridization between maize and teosinte},
  author={Ellstrand, Norman C and Garner, Lauren C and Hegde, Subray and Guadagnuolo, Roberto and Blancas, Lesley},
  journal={Journal of Heredity},
  volume={98},
  number={2},
  pages={183--187},
  year={2007},
  publisher={Am Genetic Assoc}
}

@article{gore2009first,
  title={A first-generation haplotype map of maize},
  author={Gore, Michael A and Chia, Jer-Ming and Elshire, Robert J and Sun, Qi and Ersoz, Elhan S and Hurwitz, Bonnie L and Peiffer, Jason A and McMullen, Michael D and Grills, George S and Ross-Ibarra, Jeffrey and others},
  journal={Science},
  volume={326},
  number={5956},
  pages={1115--1117},
  year={2009},
  publisher={American Association for the Advancement of Science}
}

@article{ross2009historical,
  title={Historical divergence and gene flow in the genus Zea},
  author={Ross-Ibarra, Jeffrey and Tenaillon, Maud and Gaut, Brandon S},
  journal={Genetics},
  volume={181},
  number={4},
  pages={1399--1413},
  year={2009},
  publisher={Genetics Soc America}
}

@article{van2011genetic,
  title={Genetic signals of origin, spread, and introgression in a large sample of maize landraces},
  author={van Heerwaarden, Joost and Doebley, John and Briggs, William H and Glaubitz, Jeffrey C and Goodman, Major M and Gonzalez, Jose de Jesus Sanchez and Ross-Ibarra, Jeffrey},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={3},
  pages={1088--1092},
  year={2011},
  publisher={National Acad Sciences}
}

@article{zhang2009selection,
  title={Selection on grain shattering genes and rates of rice domestication},
  author={Zhang, Lin-Bin and Zhu, Qihui and Wu, Zhi-Qiang and Ross-Ibarra, Jeffrey and Gaut, Brandon S and Ge, Song and Sang, Tao},
  journal={New Phytologist},
  volume={184},
  number={3},
  pages={708--720},
  year={2009},
  publisher={Wiley Online Library}
}

@article{matsuoka2002single,
  title={A single domestication for maize shown by multilocus microsatellite genotyping},
  author={Matsuoka, Yoshihiro and Vigouroux, Yves and Goodman, Major M and Sanchez, Jesus and Buckler, Edward and Doebley, John},
  journal={Proceedings of the National Academy of Sciences},
  volume={99},
  number={9},
  pages={6080--6084},
  year={2002},
  publisher={National Acad Sciences}
}

@article{doebley1990molecular,
  title={Molecular evidence and the evolution of maize},
  author={Doebley, John},
  journal={Economic Botany},
  volume={44},
  number={3},
  pages={6--27},
  year={1990},
  publisher={Springer}
}

@article{warburton2011gene,
  title={Gene flow among different teosinte taxa and into the domesticated maize gene pool},
  author={Warburton, Marilyn L and Wilkes, Garrison and Taba, S and Charcosset, Alain and Mir, Celine and Dumas, Fabrice and Madur, Delphine and Dreisigacker, Susanne and Bedoya, Claudia and Prasanna, BM and others},
  journal={Genetic resources and crop evolution},
  volume={58},
  number={8},
  pages={1243--1261},
  year={2011},
  publisher={Springer}
}

@article{mallet2005hybridization,
  title={Hybridization as an invasion of the genome},
  author={Mallet, James},
  journal={Trends in Ecology \& Evolution},
  volume={20},
  number={5},
  pages={229--237},
  year={2005},
  publisher={Elsevier}
}

@article{crossa1987introgression,
  title={Introgression of an exotic germplasm for improving an adapted maize population},
  author={Crossa, J and Gardner, CO},
  journal={Crop Science},
  volume={27},
  number={2},
  pages={187--190},
  year={1987},
  publisher={Crop Science Society of America}
}

@article{arnold2009adaptation,
  title={Adaptation by introgression},
  author={Arnold, Michael L and Martin, Noland H},
  journal={J Biol},
  volume={8},
  number={9},
  pages={82},
  year={2009}
}

@article{doebley1984maize,
  title={Maize introgression into teosinte-a reappraisal},
  author={Doebley, John F},
  journal={Annals of the Missouri Botanical Garden},
  pages={1100--1113},
  year={1984},
  publisher={JSTOR}
}

@article{ribaut2007marker,
  title={Marker-assisted selection to improve drought adaptation in maize: the backcross approach, perspectives, limitations, and alternatives},
  author={Ribaut, Jean-Marcel and Ragot, Michel},
  journal={Journal of Experimental Botany},
  volume={58},
  number={2},
  pages={351--360},
  year={2007},
  publisher={Soc Experiment Biol}
}

@article{hufford2013genomic,
  title={The genomic signature of crop-wild introgression in maize},
  author={Hufford, Matthew B and Lubinksy, Pesach and Pyh{\"a}j{\"a}rvi, Tanja and Devengenzo, Michael T and Ellstrand, Norman C and Ross-Ibarra, Jeffrey},
  year={2013}
}

@article{michelini1991evaluation,
  title={Evaluation of exotic and adapted maize (Zea mays L.) germplasm crosses},
  author={Michelini, Luiz Antonio},
  year={1991},
  publisher={Digital Repository@ Iowa State University, http://lib. dr. iastate. edu/}
}

@article{bernardo2009genomewide,
  title={Genomewide selection for rapid introgression of exotic germplasm in maize},
  author={Bernardo, Rex},
  journal={Crop Science},
  volume={49},
  number={2},
  pages={419--425},
  year={2009},
  publisher={Crop Science Society of America}
}

@article{jiang1999genetic,
  title={Genetic analysis of adaptation differences between highland and lowland tropical maize using molecular markers},
  author={Jiang, C and Edmeades, GO and Armstead, I and Lafitte, HR and Hayward, MD and Hoisington, D},
  journal={Theoretical and Applied Genetics},
  volume={99},
  number={7-8},
  pages={1106--1119},
  year={1999},
  publisher={Springer}
}

@article{wilkes1977hybridization,
  title={Hybridization of maize and teosinte, in Mexico and Guatemala and the improvement of maize},
  author={Wilkes, HG},
  journal={Economic Botany},
  pages={254--293},
  year={1977},
  publisher={JSTOR}
}

@article{salvi2011genetic,
  title={Genetic dissection of maize phenology using an intraspecific introgression library},
  author={Salvi, Silvio and Corneti, Simona and Bellotti, Massimo and Carraro, Nicola and Sanguineti, Maria C and Castelletti, Sara and Tuberosa, Roberto},
  journal={BMC plant biology},
  volume={11},
  number={1},
  pages={4},
  year={2011},
  publisher={BioMed Central Ltd}
}

@article{bitocchi2009introgression,
  title={Introgression from modern hybrid varieties into landrace populations of maize (Zea mays ssp. mays L.) in central Italy},
  author={Bitocchi, Elena and Nanni, Laura and Rossi, Monica and Rau, Domenico and Bellucci, Elisa and Giardini, Alessandro and Buonamici, Anna and Vendramin, Giovanni Giuseppe and Papa, Roberto},
  journal={Molecular Ecology},
  volume={18},
  number={4},
  pages={603--621},
  year={2009},
  publisher={Wiley Online Library}
}

@article{gouesnard1996evaluation,
  title={Evaluation of agronomic traits and analysis of exotic germ plasm polymorphism in adapted x exotic maize crosses},
  author={Gouesnard, B and Sanou, J and Panouille, A and Bourion, V and Boyat, A},
  journal={Theoretical and Applied Genetics},
  volume={92},
  number={3-4},
  pages={368--374},
  year={1996},
  publisher={Springer}
}

@article{placido2013introgression,
  title={Introgression of novel traits from a wild wheat relative improves drought adaptation in wheat},
  author={Placido, Dante F and Campbell, Malachy T and Folsom, Jing J and Cui, Xinping and Kruger, Greg R and Baenziger, P Stephen and Walia, Harkamal},
  journal={Plant physiology},
  volume={161},
  number={4},
  pages={1806--1819},
  year={2013},
  publisher={Am Soc Plant Biol}
}

@article{arnold2004transfer,
  title={Transfer and origin of adaptations through natural hybridization: were Anderson and Stebbins right?},
  author={Arnold, Michael L},
  journal={The Plant Cell},
  volume={16},
  number={3},
  pages={562--570},
  year={2004},
  publisher={Am Soc Plant Biol}
}

@article{wilkes1970teosinte,
  title={Teosinte introgression in the maize of the Nobogame Valley},
  author={Wilkes, H Garrison},
  journal={Botanical Museum Leaflets, Harvard University},
  pages={297--311},
  year={1970},
  publisher={JSTOR}
}

@article{lam2010resequencing,
  title={Resequencing of 31 wild and cultivated soybean genomes identifies patterns of genetic diversity and selection},
  author={Lam, Hon-Ming and Xu, Xun and Liu, Xin and Chen, Wenbin and Yang, Guohua and Wong, Fuk-Ling and Li, Man-Wah and He, Weiming and Qin, Nan and Wang, Bo and others},
  journal={Nature genetics},
  volume={42},
  number={12},
  pages={1053--1059},
  year={2010},
  publisher={Nature Publishing Group}
}

@article{kato2002introgression,
  title={Introgression of chromosome knobs from Zea diploperennis into maize},
  author={KATO Y, TA and SANCHEZ G, JJ},
  journal={Maydica},
  volume={47},
  number={1},
  pages={33--50},
  year={2002},
  publisher={Maydica}
}

@article{fukunaga2005genetic,
  title={Genetic diversity and population structure of teosinte},
  author={Fukunaga, Kenji and Hill, Jason and Vigouroux, Yves and Matsuoka, Yoshihiro and Sanchez, Jesus and Liu, Kejun and Buckler, Edward S and Doebley, John},
  journal={Genetics},
  volume={169},
  number={4},
  pages={2241--2254},
  year={2005},
  publisher={Genetics Soc America}
}

@article{reagon2006cultivated,
  title={Cultivated Helianthus annuus (Asteraceae) volunteers as a genetic “bridge” to weedy sunflower populations in North America},
  author={Reagon, Michael and Snow, Allison A},
  journal={American Journal of Botany},
  volume={93},
  number={1},
  pages={127--133},
  year={2006},
  publisher={Botanical Soc America}
}

@article{broyles2002hybrid,
  title={Hybrid bridges to gene flow: a case study in milkweeds (Asclepias)},
  author={Broyles, Steven B},
  journal={Evolution},
  volume={56},
  number={10},
  pages={1943--1953},
  year={2002},
  publisher={Wiley Online Library}
}

@article{chapman2006letting,
  title={Letting the gene out of the bottle: the population genetics of genetically modified crops},
  author={Chapman, Mark A and Burke, John M},
  journal={New Phytologist},
  volume={170},
  number={3},
  pages={429--443},
  year={2006},
  publisher={Wiley Online Library}
}

@article{collins1921dominance,
  title={Dominance and the vigor of first generation hybrids},
  author={Collins, GN},
  journal={American Naturalist},
  pages={116--133},
  year={1921},
  publisher={JSTOR}
}

@article{doebley1987patterns,
  title={Patterns of isozyme variation between maize and Mexican annual teosinte},
  author={Doebley, John and Goodman, Major M and Stuber, Charles W},
  journal={Economic Botany},
  volume={41},
  number={2},
  pages={234--246},
  year={1987},
  publisher={Springer}
}

@article{slatkin1987gene,
  title={Gene flow and the geographic structure of natural populations},
  author={Slatkin, Montgomery},
  journal={Science},
  volume={236},
  number={4803},
  pages={787--792},
  year={1987},
  publisher={American Association for the Advancement of Science}
}

@article{zohary1969wild,
  title={The wild diploid progenitors of wheat and their breeding value},
  author={Zohary, Daniel and Harlan, Jack R and Vardi, A},
  journal={Euphytica},
  volume={18},
  number={1},
  pages={58--65},
  year={1969},
  publisher={Springer}
}

@article{von2006ab,
  title={AB-QTL analysis in spring barley: II. Detection of favourable exotic alleles for agronomic traits introgressed from wild barley (H. vulgare ssp. spontaneum)},
  author={Von Korff, M and Wang, H and L{\'e}on, J and Pillen, K},
  journal={Theoretical and Applied Genetics},
  volume={112},
  number={7},
  pages={1221--1231},
  year={2006},
  publisher={Springer}
}

@article{nagy1997local,
  title={Local adaptation in two subspecies of an annual plant: implications for migration and gene flow},
  author={Nagy, Eric S and Rice, Kevin J},
  journal={Evolution},
  pages={1079--1089},
  year={1997},
  publisher={JSTOR}
}

@article{abbott2013hybridization,
  title={Hybridization and speciation},
  author={Abbott, R and Albach, D and Ansell, S and Arntzen, JW and Baird, SJE and Bierne, N and Boughman, J and Brelsford, A and Buerkle, CA and Buggs, R and others},
  journal={Journal of Evolutionary Biology},
  volume={26},
  number={2},
  pages={229--246},
  year={2013},
  publisher={Wiley Online Library}
}

@article{nevo2010drought,
  title={Drought and salt tolerances in wild relatives for wheat and barley improvement},
  author={Nevo, Eviatar and Chen, Guoxiong},
  journal={Plant, cell \& environment},
  volume={33},
  number={4},
  pages={670--685},
  year={2010},
  publisher={Wiley Online Library}
}

@article{hussain1997development,
  title={Development of a fertile genetic bridge between Trifolium ambiguum M. Bieb. and T. repens L.},
  author={Hussain, SW and Williams, WM},
  journal={Theoretical and applied genetics},
  volume={95},
  number={4},
  pages={678--690},
  year={1997},
  publisher={Springer}
}

@article{hegde2004hybridization,
  title={Hybridization and introgression between bread wheat and wild and weedy relatives in North America},
  author={Hegde, SG and Waines, JG},
  journal={Crop Science},
  volume={44},
  number={4},
  pages={1145--1155},
  year={2004},
  publisher={Crop Science Society of America}
}

@article{autrique1995molecular,
  title={Molecular markers for four leaf rust resistance genes introgressed into wheat from wild relatives},
  author={Autrique, Enrique and Tanksley, Steven D and Sorrells, Mark E and Singh, Ravi P},
  journal={Genome},
  volume={38},
  number={1},
  pages={75--83},
  year={1995},
  publisher={NRC Research Press}
}

@article{Feuillet200824,
title = "Cereal breeding takes a walk on the wild side ",
journal = "Trends in Genetics ",
volume = "24",
number = "1",
pages = "24 - 32",
year = "2008",
note = "",
issn = "0168-9525",
doi = "http://dx.doi.org/10.1016/j.tig.2007.11.001",
url = "http://www.sciencedirect.com/science/article/pii/S0168952507003484",
author = "Catherine Feuillet and Peter Langridge and Robbie Waugh"
}

@article{able2007capturing,
  title={Capturing diversity in the cereals: many options but little promiscuity},
  author={Able, Jason A and Langridge, Peter and Milligan, Andrew S},
  journal={Trends in plant science},
  volume={12},
  number={2},
  pages={71--79},
  year={2007},
  publisher={Elsevier}
}

@article{varshney2006advances,
  title={Advances in cereal genomics and applications in crop breeding},
  author={Varshney, Rajeev K and Hoisington, David A and Tyagi, Akhilesh K},
  journal={Trends in biotechnology},
  volume={24},
  number={11},
  pages={490--499},
  year={2006},
  publisher={Elsevier}
}

@article{brown2009complex,
  title={The complex origins of domesticated crops in the Fertile Crescent},
  author={Brown, Terence A and Jones, Martin K and Powell, Wayne and Allaby, Robin G},
  journal={Trends in Ecology \& Evolution},
  volume={24},
  number={2},
  pages={103--109},
  year={2009},
  publisher={Elsevier}
}

@article{suneson1969dynamic,
  title={A dynamic population of weedy rye},
  author={Suneson, CA and Rachie, KO and Khush, GS},
  journal={Crop Science},
  volume={9},
  number={2},
  pages={121--124},
  year={1969},
  publisher={Crop Science Society of America}
}

@article{stewart2003transgene,
  title={Transgene introgression from genetically modified crops to their wild relatives},
  author={Stewart, C Neal and Halfhill, Matthew D and Warwick, Suzanne I},
  journal={Nature Reviews Genetics},
  volume={4},
  number={10},
  pages={806--817},
  year={2003},
  publisher={Nature Publishing Group}
}

@article{zamir2001improving,
  title={Improving plant breeding with exotic genetic libraries},
  author={Zamir, Dani},
  journal={Nature reviews genetics},
  volume={2},
  number={12},
  pages={983--989},
  year={2001},
  publisher={Nature Publishing Group}
}

@article{tanksley1997seed,
  title={Seed banks and molecular maps: unlocking genetic potential from the wild},
  author={Tanksley, Steven D and McCouch, Susan R},
  journal={Science},
  volume={277},
  number={5329},
  pages={1063--1066},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{xie2008wild,
  title={Wild emmer: genetic resources, gene mapping and potential for wheat improvement},
  author={Xie, Weilong and Nevo, Eviatar},
  journal={Euphytica},
  volume={164},
  number={3},
  pages={603--614},
  year={2008},
  publisher={Springer}
}

@book{grant1950genetic,
  title={Genetic and taxonomic studies in Gilia},
  author={Grant, Verne and Grant, Karen A},
  year={1950},
  publisher={Rancho Santa Ana Botanic Garden}
}

@article{hajjar2007use,
  title={The use of wild relatives in crop improvement: a survey of developments over the last 20 years},
  author={Hajjar, Reem and Hodgkin, Toby},
  journal={Euphytica},
  volume={156},
  number={1-2},
  pages={1--13},
  year={2007},
  publisher={Springer}
}

@article{jorgensen1992discovery,
  title={Discovery, characterization and exploitation of Mlo powdery mildew resistance in barley},
  author={J{\o}rgensen, I Helms},
  journal={Euphytica},
  volume={63},
  number={1-2},
  pages={141--152},
  year={1992},
  publisher={Springer}
}

@article{gepts2001origins,
  title={Origins of plant agriculture and major crop plants},
  author={Gepts, Paul},
  journal={Our fragile world: Challenges and opportunities for sustainable development. Oxford: EOLSS Publishers},
  pages={629--637},
  year={2001}
}

@article{smith1997initial,
  title={The initial domestication of Cucurbita pepo in the Americas 10,000 years ago},
  author={Smith, Bruce D},
  journal={Science},
  volume={276},
  number={5314},
  pages={932--934},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{piperno2001earliest,
  title={The earliest archaeological maize (Zea mays L.) from highland Mexico: new accelerator mass spectrometry dates and their implications},
  author={Piperno, Dolores R and Flannery, Kent V},
  journal={Proceedings of the National Academy of Sciences},
  volume={98},
  number={4},
  pages={2101--2103},
  year={2001},
  publisher={National Acad Sciences}
}

@article{colledge2004archaeobotanical,
  title={Archaeobotanical Evidence for the Spread of Farming in the Eastern Mediterranean 1},
  author={Colledge, Sue and Conolly, James and Shennan, Stephen and Bellwood, Peter and Bouby, Laurent and Hansen, Julie and Harris, DavidR and Kotsakis, Konstantinos and zdoan, Mehmet and Peltenburg, Edgar and others},
  journal={Current anthropology},
  volume={45},
  number={S4},
  pages={S35--S58},
  year={2004},
  publisher={JSTOR}
}

@article{zhijun1998middle,
  title={The Middle Yangtze region in China is one place where rice was domesticated: phytolith evidence from the Diaotonghuan Cave, Northern Jiangxi},
  author={Zhijun, Zhao},
  journal={Antiquity},
  volume={72},
  number={278},
  pages={885--897},
  year={1998},
  publisher={Cambridge Univ Press}
}

@article{hart2002age,
  title={The age of the common bean (Phaseolus vulgaris L.) in the northern Eastern Woodlands of North America},
  author={Hart, John P and Asch, David L and Scarry, C Margaret and Crawford, Gary W},
  journal={Antiquity},
  volume={76},
  number={292},
  pages={377--385},
  year={2002},
  publisher={Cambridge Univ Press}
}

@article{crites1993domesticated,
  title={Domesticated sunflower in fifth millennium BP temporal context: new evidence from Middle Tennessee},
  author={Crites, Gary D},
  journal={American Antiquity},
  pages={146--148},
  year={1993},
  publisher={JSTOR}
}

@article{piperno2000starch,
  title={Starch grains reveal early root crop horticulture in the Panamanian tropical forest},
  author={Piperno, Dolores R and Ranere, Anthony J and Holst, Irene and Hansell, Patricia},
  journal={Nature},
  volume={407},
  number={6806},
  pages={894--897},
  year={2000},
  publisher={Nature Publishing Group}
}

@article{pope2001origin,
  title={Origin and environmental setting of ancient agriculture in the lowlands of Mesoamerica},
  author={Pope, Kevin O and Pohl, Mary ED and Jones, John G and Lentz, David L and Von Nagy, Christopher and Vega, Francisco J and Quitmyer, Irvy R},
  journal={Science},
  volume={292},
  number={5520},
  pages={1370--1373},
  year={2001},
  publisher={American Association for the Advancement of Science}
}

@inproceedings{smith1989origins,
  title={Origins of agriculture in eastern North America},
  author={Smith, Bruce David},
  year={1989},
  organization={American Association for the Advancement of Science}
}

@book{ammerman2014neolithic,
  title={The Neolithic transition and the genetics of populations in Europe},
  author={Ammerman, Albert J and Cavalli-Sforza, Luigi Luca},
  year={2014},
  publisher={Princeton University Press}
}

@article{hufford2012comparative,
  title={Comparative population genomics of maize domestication and improvement},
  author={Hufford, Matthew B and Xu, Xun and Van Heerwaarden, Joost and Pyh{\"a}j{\"a}rvi, Tanja and Chia, Jer-Ming and Cartwright, Reed A and Elshire, Robert J and Glaubitz, Jeffrey C and Guill, Kate E and Kaeppler, Shawn M and others},
  journal={Nature genetics},
  volume={44},
  number={7},
  pages={808--811},
  year={2012},
  publisher={Nature Publishing Group}
}

@book{lumholtz1902unknown,
  title={Unknown Mexico: a record of five years' exploration among the tribes of the western Sierra Madre; in the Tierra Caliente of Tepic and Jalisco; and among the Tarascos of Michoacan},
  author={Lumholtz, Carl},
  volume={2},
  year={1902},
  publisher={C. Scribner's sons}
}

@book{smith1995emergence,
  title={The emergence of agriculture},
  author={Smith, Bruce D},
  year={1995},
  publisher={Scientific American Library New York}
}

@article{heun1997site,
  title={Site of einkorn wheat domestication identified by DNA fingerprinting},
  author={Heun, Manfred and Sch{\"a}fer-Pregl, Ralf and Klawan, Dieter and Castagna, Renato and Accerbi, Monica and Borghi, Basilio and Salamini, Francesco},
  journal={Science},
  volume={278},
  number={5341},
  pages={1312--1314},
  year={1997},
  publisher={American Association for the Advancement of Science}
}

@article{nesbitt1998wheat,
  title={Wheat domestication: archaeobotanical evidence},
  author={Nesbitt, Mark and Samuel, Delwen},
  journal={Science},
  volume={279},
  number={5356},
  pages={1431--1431},
  year={1998}
}

@article{lev2000cradle,
  title={The cradle of agriculture},
  author={Lev-Yadun, Simcha and Gopher, Avi and Abbo, Shahal},
  journal={Science},
  volume={288},
  number={5471},
  pages={1602},
  year={2000},
  publisher={The American Association for the Advancement of Science}
}

@book{zohary2012domestication,
  title={Domestication of Plants in the Old World: The origin and spread of domesticated plants in Southwest Asia, Europe, and the Mediterranean Basin},
  author={Zohary, Daniel and Hopf, Maria and Weiss, Ehud},
  year={2012},
  publisher={Oxford University Press on Demand}
}

@article{ozkan2005reconsideration,
  title={A reconsideration of the domestication geography of tetraploid wheats},
  author={Ozkan, Hakan and Brandolini, Andrea and Pozzi, Carlo and Effgen, Sieglinde and Wunder, J and Salamini, Francesco},
  journal={Theoretical and Applied Genetics},
  volume={110},
  number={6},
  pages={1052--1060},
  year={2005},
  publisher={Springer}
}

@article {TPJ:TPJ2772,
author = {Yahiaoui, Nabila and Brunner, Susanne and Keller, Beat},
title = {Rapid generation of new powdery mildew resistance genes after wheat domestication},
journal = {The Plant Journal},
volume = {47},
number = {1},
publisher = {Blackwell Publishing Ltd},
issn = {1365-313X},
url = {http://dx.doi.org/10.1111/j.1365-313X.2006.02772.x},
doi = {10.1111/j.1365-313X.2006.02772.x},
pages = {85--98},
keywords = {disease resistance, evolution, wheat, powdery mildew, arms race, domestication},
year = {2006},
}

@article{tanno2006fast,
  title={How fast was wild wheat domesticated?},
  author={Tanno, Ken-ichi and Willcox, George},
  journal={Science},
  volume={311},
  number={5769},
  pages={1886--1886},
  year={2006},
  publisher={American Association for the Advancement of Science}
}

@article{kilian2006haplotype,
  title={Haplotype structure at seven barley genes: relevance to gene pool bottlenecks, phylogeny of ear type and site of barley domestication},
  author={Kilian, Benjamin and {\"O}zkan, Hakan and Kohl, Jochen and von Haeseler, Arndt and Barale, Francesca and Deusch, Oliver and Brandolini, Andrea and Yucel, Cemal and Martin, William and Salamini, Francesco},
  journal={Molecular Genetics and Genomics},
  volume={276},
  number={3},
  pages={230--241},
  year={2006},
  publisher={Springer}
}

@article{badr2000origin,
  title={On the origin and domestication history of barley (Hordeum vulgare)},
  author={Badr, Abdelfattah and Sch, R and El Rabey, H and Effgen, S and Ibrahim, HH and Pozzi, C and Rohde, W and Salamini, F and others},
  journal={Molecular Biology and Evolution},
  volume={17},
  number={4},
  pages={499--510},
  year={2000},
  publisher={SMBE}
}

@article{azhaguvel2007phylogenetic,
  title={A phylogenetic analysis based on nucleotide sequence of a marker linked to the brittle rachis locus indicates a diphyletic origin of barley},
  author={Azhaguvel, Perumal and Komatsuda, Takao},
  journal={Annals of botany},
  volume={100},
  number={5},
  pages={1009--1015},
  year={2007},
  publisher={Annals Botany Co}
}

@article{von1995ecogeographical,
  title={An ecogeographical study of the genus Hordeum. Systematic and ecogeographic studies on crop genepools 7},
  author={von Bothmer, R von and Jacobsen, N and Baden, C and J{\o}rgensen, RB and Linde-Laursen, I},
  journal={International Plant Genetic Resources Institute, Rome},
  year={1995}
}

@article{hanson2000mapping,
  title={Mapping a wild tomato introgression associated with tomato yellow leaf curl virus resistance in a cultivated tomato line},
  author={Hanson, Peter M and Bernacchi, Dario and Green, Sylvia and Tanksley, Steven D and Muniyappa, Venkataramappa and Padmaja, Attiganal S and Chen, Huei-mei and Kuo, George and Fang, Denise and Chen, Jen-tzu},
  journal={Journal of the American Society for Horticultural Science},
  volume={125},
  number={1},
  pages={15--20},
  year={2000},
  publisher={American Society for Horticultural Science}
}

@article{banerjee1990transfer,
  title={Transfer of tomato leaf curl virus resistance from Lycopersicon hirsutum f. glabratum to L. esculentum},
  author={Banerjee, MK and others},
  journal={Plant Breeding},
  volume={105},
  number={2},
  pages={156--159},
  year={1990},
  publisher={Wiley Online Library}
}

@inproceedings{sims1979history,
  title={History of tomato production for industry around the world},
  author={Sims, WL},
  booktitle={Symposium on Production of Tomatoes for Processing 100},
  pages={25--26},
  year={1979}
}

@article{rick1988tomato,
  title={Tomato-like nightshades: affinities, autoecology, and breeders’ opportunities},
  author={Rick, Charles M},
  journal={Economic Botany},
  volume={42},
  number={2},
  pages={145--154},
  year={1988},
  publisher={Springer}
}

@article{miller1990rflp,
  title={RFLP analysis of phylogenetic relationships and genetic variation in the genus Lycopersicon},
  author={Miller, JC and Tanksley, SD},
  journal={Theoretical and applied genetics},
  volume={80},
  number={4},
  pages={437--448},
  year={1990},
  publisher={Springer}
}

@article{rick1982potential,
  title={The potential of exotic germplasm for tomato improvement},
  author={Rick, CM},
  year={1982},
  publisher={Academic Press}
}

@article{li2006rice,
  title={Rice domestication by reducing shattering},
  author={Li, Changbao and Zhou, Ailing and Sang, Tao},
  journal={science},
  volume={311},
  number={5769},
  pages={1936--1939},
  year={2006},
  publisher={American Association for the Advancement of Science}
}

@book{oka2012origin,
  title={Origin of cultivated rice},
  author={Oka, H-I},
  year={2012},
  publisher={Elsevier}
}

@incollection{brar1997alien,
  title={Alien introgression in rice},
  author={Brar, DS and Khush, GS},
  booktitle={Oryza: From Molecule to Plant},
  pages={35--47},
  year={1997},
  publisher={Springer}
}

@article{khush1974inheritance,
  title={Inheritance of resistance to grassy stunt virus and its vector in rice},
  author={KHUSH, GURDEV S and Ling, KC},
  journal={Journal of Heredity},
  volume={65},
  number={3},
  pages={135--136},
  year={1974},
  publisher={Am Genetic Assoc}
}

@article{spooner2005single,
  title={A single domestication for potato based on multilocus amplified fragment length polymorphism genotyping},
  author={Spooner, David M and McLean, Karen and Ramsay, Gavin and Waugh, Robbie and Bryan, Glenn J},
  journal={Proceedings of the National Academy of Sciences of the United States of America},
  volume={102},
  number={41},
  pages={14694--14699},
  year={2005},
  publisher={National Acad Sciences}
}

@article{second1982origin,
  title={Origin of the genic diversity of cultivated rice (Oryza spp.): study of the polymorphism scored at 40 isozyme loci.},
  author={Second, Gerard},
  journal={遺伝学雑誌},
  volume={57},
  number={1},
  pages={25--57},
  year={1982},
  publisher={The Genetics Society of Japan}
}

@article{van1999tight,
  title={Tight physical linkage of the nematode resistance gene Gpa2 and the virus resistance gene Rx on a single segment introgressed from the wild species Solanum tuberosum subsp. andigena CPC 1673 into cultivated potato},
  author={van der Voort, Jeroen Rouppe and Kanyuka, Konstantin and van der Vossen, Edwin and Bendahmane, Abdelhafid and Mooijman, Paul and Klein-Lankhorst, Ren{\'e} and Stiekema, Willem and Baulcombe, David and Bakker, Jaap},
  journal={Molecular plant-microbe interactions},
  volume={12},
  number={3},
  pages={197--206},
  year={1999},
  publisher={Am Phytopath Society}
}

@article{van2003ancient,
  title={An ancient R gene from the wild potato species Solanum bulbocastanum confers broad-spectrum resistance to Phytophthora infestans in cultivated potato and tomato},
  author={Van Der Vossen, Edwin and Sikkema, Anne and Hekkert, Bas te Lintel and Gros, Jack and Stevens, Patricia and Muskens, Marielle and Wouters, Doret and Pereira, Andy and Stiekema, Willem and Allefs, Sjefke},
  journal={The Plant Journal},
  volume={36},
  number={6},
  pages={867--882},
  year={2003},
  publisher={Wiley Online Library}
}

@article{johnston1980significance,
  title={The significance of genic balance to endosperm development in interspecific crosses},
  author={Johnston, SA and Den Nijs, TPM and Peloquin, SJ and Hanneman Jr, RE},
  journal={Theoretical and applied genetics},
  volume={57},
  number={1},
  pages={5--9},
  year={1980},
  publisher={Springer}
}

@article{jarvis1999wild,
  title={Wild relatives and crop cultivars: detecting natural introgression and farmer selection of new genetic combinations in agroecosystems},
  author={Jarvis, Devra I and Hodgkin, Toby},
  journal={Molecular ecology},
  volume={8},
  number={s1},
  pages={S159--S173},
  year={1999},
  publisher={Wiley Online Library}
}

@article{rieseberg1999introgression,
  title={Introgression between the cultivated sunflower and a sympatric wild relative, Helianthus petiolaris (Asteraceae)},
  author={Rieseberg, Loren H and Kim, Min Ju and Seiler, Gerald J},
  journal={International Journal of Plant Sciences},
  volume={160},
  number={1},
  pages={102--108},
  year={1999},
  publisher={JSTOR}
}

@article{kim1999genetic,
  title={Genetic architecture of species differences in annual sunflowers: implications for adaptive trait introgression},
  author={Kim, Seung-Chul and Rieseberg, Loren H},
  journal={Genetics},
  volume={153},
  number={2},
  pages={965--977},
  year={1999},
  publisher={Genetics Soc America}
}

@article{papa2003asymmetry,
  title={Asymmetry of gene flow and differential geographical structure of molecular diversity in wild and domesticated common bean (Phaseolus vulgaris L.) from Mesoamerica},
  author={Papa, R and Gepts, P},
  journal={Theoretical and Applied Genetics},
  volume={106},
  number={2},
  pages={239--250},
  year={2003},
  publisher={Springer}
}

@article{blackman2011sunflower,
  title={Sunflower domestication alleles support single domestication center in eastern North America},
  author={Blackman, Benjamin K and Scascitelli, Moira and Kane, Nolan C and Luton, Harry H and Rasmussen, David A and Bye, Robert A and Lentz, David L and Rieseberg, Loren H},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={34},
  pages={14360--14365},
  year={2011},
  publisher={National Acad Sciences}
}

@article{wills2006chloroplast,
  title={Chloroplast DNA variation confirms a single origin of domesticated sunflower (Helianthus annuus L.)},
  author={Wills, David M and Burke, John M},
  journal={Journal of Heredity},
  volume={97},
  number={4},
  pages={403--408},
  year={2006},
  publisher={Am Genetic Assoc}
}

@article{vandenberg1998collapse,
  title={Collapse of morphological species in the wild potato Solanum brevicaule complex (Solanaceae: sect. Petota).},
  author={Vandenberg, R and Miller, J and Ugarte, M and Kardolus, J and Villand, Julie and Nienhuis, James and Spooner, D},
  journal={American Journal of Botany},
  volume={85},
  number={1},
  pages={92--92},
  year={1998},
  publisher={Botanical Soc America}
}

@book{hawkes1990potato,
  title={The potato: evolution, biodiversity and genetic resources.},
  author={Hawkes, John Gregory and others},
  year={1990},
  publisher={Belhaven Press}
}

@article{spooner2001potato,
  title={Potato systematics and germplasm collecting, 1989--2000},
  author={Spooner, David M and Hijmans, Robert J},
  journal={American Journal of Potato Research},
  volume={78},
  number={4},
  pages={237--268},
  year={2001},
  publisher={Springer}
}

@article{sharma2000origin,
  title={Origin of O. sativa and its ecotypes},
  author={Sharma, SD and Tripathy, Smita and Biswal, Jyostnamayee},
  journal={Rice breeding and genetics: research priorities and challenges. Enfield, NH, USA: Science Publications},
  pages={349--369},
  year={2000}
}

@article{oka1988origin,
  title={Origin of cultivated rice. Jpn},
  author={Oka, HI},
  journal={Sci. Soc. Press, Elsevier, Tokyo},
  volume={254},
  year={1988}
}

@article{robertson2007genetic,
  title={Genetic resources of tomato (Lycopersicon esculentum Mill.) and wild relatives},
  author={Robertson, LARRY D and Labate, JOANNE A and Razdan, MK and Mattoo, AK},
  journal={Genetic Improvement of Solanaceous Crops. Tomato},
  volume={2},
  pages={25--75},
  year={2007}
}

@article{giles1985progenitor,
  title={The progenitor of barley (Hordeum vulgare spp. spontaneum)-its importance as a gene resource [incl. hordeins, Erysiphe graminis].},
  author={Giles, BE and others},
  journal={Sveriges Utsaedesfoerenings Tidskrift},
  year={1985}
}

@article{bai2007domestication,
  title={Domestication and breeding of tomatoes: what have we gained and what can we gain in the future?},
  author={Bai, Yuling and Lindhout, Pim},
  journal={Annals of Botany},
  volume={100},
  number={5},
  pages={1085--1094},
  year={2007},
  publisher={Annals Botany Co}
}

@article{rick1953novel,
  title={Novel variation in tomato species hybrids},
  author={Rick, Charles M and Smith, Paul G},
  journal={American Naturalist},
  pages={359--373},
  year={1953},
  publisher={JSTOR}
}

http://www.landscapeimagery.com/tomato.html
Cox S. 2000. I Say Tomayto, You Say Tomahto. http://www.landscapeimagery.com/tomato.html

@article{nesbitt2002comparative,
  title={Comparative sequencing in the genus Lycopersicon: implications for the evolution of fruit size in the domestication of cultivated tomatoes},
  author={Nesbitt, T Clint and Tanksley, Steven D},
  journal={Genetics},
  volume={162},
  number={1},
  pages={365--379},
  year={2002},
  publisher={Genetics Soc America}
}

@article{heiser1951hybridization,
  title={Hybridization in the annual sunflowers: Helianthus annuus$\times$ H. debilis var. cucumerifolius},
  author={Heiser Jr, Charles B},
  journal={Evolution},
  pages={42--51},
  year={1951},
  publisher={JSTOR}
}

@article{rieseberg1999hybrid,
  title={Hybrid zones and the genetic architecture of a barrier to gene flow between two sunflower species},
  author={Rieseberg, Loren H and Whitton, Jeannette and Gardner, Keith},
  journal={Genetics},
  volume={152},
  number={2},
  pages={713--727},
  year={1999},
  publisher={Genetics Soc America}
}

@article{kwak2009putative,
  title={The Putative Mesoamerican Domestication Center of Is Located in the Lerma--Santiago Basin of Mexico},
  author={Kwak, Myounghai and Kami, James A and Gepts, Paul},
  journal={Crop Science},
  volume={49},
  number={2},
  pages={554--563},
  year={2009},
  publisher={Crop Science Society of America}
}

@article{kami1995identification,
  title={Identification of presumed ancestral DNA sequences of phaseolin in Phaseolus vulgaris},
  author={Kami, James and Vel{\'a}squez, V Becerra and Debouck, Daniel G and Gepts, Paul},
  journal={Proceedings of the National Academy of Sciences},
  volume={92},
  number={4},
  pages={1101--1104},
  year={1995},
  publisher={National Acad Sciences}
}

@article{park1987transfer,
  title={Transfer of common blight (Xanthomonas campestris pv phaseoli) resistance from Phaseolus coccineus Lam. to P. vulgaris L. through interspecific hybridization},
  author={Park, SJ and Dhanvantari, BN},
  journal={Canadian journal of plant science},
  volume={67},
  number={3},
  pages={685--695},
  year={1987},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{schwartz2006inheritance,
  title={Inheritance of white mold resistance in Phaseolus vulgaris$\times$ P. coccineus crosses},
  author={Schwartz, Howard F and Otto, Kristen and Ter{\'a}n, Henry and Lema, Margarita and Singh, Shree P},
  journal={Plant Disease},
  volume={90},
  number={9},
  pages={1167--1170},
  year={2006},
  publisher={Am Phytopath Society}
}

@article{voysest1991bean,
  title={Bean cultivars: classes and commercial seed types},
  author={Voysest, Oswaldo and Dessert, Michael},
  journal={Common beans: research for crop improvement},
  pages={119--162},
  year={1991},
  publisher={CAB Int Wallingford, UK and CIAT, Cali, Colombia}
}

@article{gutierrez1985heterosis,
  title={Heterosis and inbreeding depression in dry bush beans, Phaseolus vulgaris L.},
  author={GUTIERREZ, J ARIEL and SINGH, SHREE P},
  journal={Canadian journal of plant science},
  volume={65},
  number={2},
  pages={243--249},
  year={1985},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{singh1995potential,
  title={Potential of wild common bean for seed yield improvement of cultivars in the tropics},
  author={Singh, Shree P and Molina, Albeiro and Gepts, Paul},
  journal={Canadian Journal of Plant Science},
  volume={75},
  number={4},
  pages={807--813},
  year={1995},
  publisher={NRC Research Press Ottawa, Canada}
}

@article{burle2010microsatellite,
  title={Microsatellite diversity and genetic structure among common bean (Phaseolus vulgaris L.) landraces in Brazil, a secondary center of diversity},
  author={Burle, Mar{\'\i}lia Lobo and Fonseca, Jaime Roberto and Kami, James A and Gepts, Paul},
  journal={Theoretical and Applied Genetics},
  volume={121},
  number={5},
  pages={801--813},
  year={2010},
  publisher={Springer}
}

@article{ramalho2006,
[[Ramalho MAP, Abreu AFB (2006) Cultivares. In: Vieira C et al (eds) Feijao, 2nd edn. UFV, Vicosa, pp415-436]]

@article{gepts1986phaseolin,
  title={Phaseolin variability among wild and cultivated common beans (Phaseolus vulgaris) from Colombia},
  author={Gepts, P and Bliss, FA},
  journal={Economic Botany},
  volume={40},
  number={4},
  pages={469--478},
  year={1986},
  publisher={Springer}
}

@article{paredes1995extensive,
  title={Extensive introgression of Middle American germplasm into Chilean common bean cultivars},
  author={Paredes, Oscar M and Gepts, Paul},
  journal={Genetic Resources and Crop Evolution},
  volume={42},
  number={1},
  pages={29--41},
  year={1995},
  publisher={Springer}
}

@article{sotelo1995comparative,
  title={Comparative study of the chemical composition of wild and cultivated beans (Phaseolus vulgaris)},
  author={Sotelo, Angela and Sousa, H and Sanchez, M},
  journal={Plant Foods for Human Nutrition},
  volume={47},
  number={2},
  pages={93--100},
  year={1995},
  publisher={Springer}
}

@inproceedings{debouck1988implications,
  title={Implications for bean breeders of studies on the origin of common beans, Phaseolus vulgaris L},
  author={Debouck, DG and THOME, S},
  booktitle={THE BEAN breeders. Cali, Workshop at CIAT},
  year={1988}
}

@article{debouck1993genetic,
  title={Genetic diversity and ecological distribution ofPhaseolus vulgaris (Fabaceae) in northwestern South America},
  author={Debouck, Daniel G and Toro, Orlando and Paredes, Oscar M and Johnson, William C and Gepts, Paul},
  journal={Economic botany},
  volume={47},
  number={4},
  pages={408--423},
  year={1993},
  publisher={Springer}
}

@article{kornegay1991inheritance,
  title={Inheritance of resistance to Acanthoscelides obtectus in a wild common bean accession crossed to commercial bean cultivars},
  author={Kornegay, Julia L and Cardona, Cesar},
  journal={Euphytica},
  volume={52},
  number={2},
  pages={103--111},
  year={1991},
  publisher={Springer}
}

@article{komatsuda2007six,
  title={Six-rowed barley originated from a mutation in a homeodomain-leucine zipper I-class homeobox gene},
  author={Komatsuda, Takao and Pourkheirandish, Mohammad and He, Congfen and Azhaguvel, Perumal and Kanamori, Hiroyuki and Perovic, Dragan and Stein, Nils and Graner, Andreas and Wicker, Thomas and Tagiri, Akemi and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={104},
  number={4},
  pages={1424--1429},
  year={2007},
  publisher={National Acad Sciences}
}

@article{harlan1968origin,
  title={On the origin of barley},
  author={Harlan, Jack R and others},
  journal={USDA Agriculture Handbook},
  volume={338},
  pages={9--31},
  year={1968}
}

@misc{von1995ecographical,
  title={An ecographical study of genus Hordeum},
  author={Von Bothmer, R and Jacobsen, N and Baden, C and J{\o}rgensen, RB and Linde-Laursen, I},
  year={1995},
  publisher={Systematic and ecographic studies on crop gene pools}
}

@article{ellis2000wild,
  title={Wild barley: a source of genes for crop improvement in the 21st century?},
  author={Ellis, RP and Forster, BP and Robinson, D and Handley, LL and Gordon, DC and Russell, JR and Powell, W},
  journal={Journal of experimental botany},
  volume={51},
  number={342},
  pages={9--17},
  year={2000},
  publisher={Soc Experiment Biol}
}

@article{handley1994chromosome,
  title={Chromosome 4 controls potential water use efficiency ($\delta$13C) in barley},
  author={Handley, Linda L and Nevo, Eviatar and Raven, John A and MartInez-Carrasco, Rafael and Scrimgeour, Charles M and Pakniyat, Hassan and Forster, Brian P},
  journal={Journal of Experimental Botany},
  volume={45},
  number={11},
  pages={1661--1663},
  year={1994},
  publisher={Soc Experiment Biol}
}

@article{brown1978outcrossing,
  title={Outcrossing rates and heterozygosity in natural populations of Hordeum spontaneum Koch in Israel},
  author={Brown, AHD and Zohary, D and Nevo, E},
  journal={Heredity},
  volume={41},
  number={1},
  pages={49--62},
  year={1978}
}

@article{takahashi1955origin,
  title={The origin and evolution of cultivated barley},
  author={Takahashi, Ryuhei},
  journal={Advances in genetics},
  volume={7},
  pages={227--266},
  year={1955},
  publisher={Elsevier}
}

@article{huaman2002reclassification,
  title={Reclassification of landrace populations of cultivated potatoes (Solanum sect. Petota)},
  author={Huam{\'a}n, Z{\'o}simo and Spooner, David M},
  journal={American Journal of Botany},
  volume={89},
  number={6},
  pages={947--965},
  year={2002},
  publisher={Botanical Soc America}
}

@article{harlan1966distribution,
  title={Distribution of wild wheats and barley},
  author={Harlan, Jack R and Zohary, Daniel},
  journal={Science},
  volume={153},
  number={3740},
  pages={1074--1080},
  year={1966},
  publisher={American Association for the Advancement of Science}
}

@article{correll1962potato,
  title={The potato and its wild relatives.},
  author={Correll, Donovan Stewart and others},
  journal={The potato and its wild relatives.},
  year={1962},
  publisher={Texas Research Foundation, Renner.}
}

@article{dussle2004pl,
  title={Pl Arg from Helianthus argophyllus is unlinked to other known downy mildew resistance genes in sunflower},
  author={Dussle, CM and Hahn, V and Knapp, SJ and Bauer, E},
  journal={Theoretical and applied genetics},
  volume={109},
  number={5},
  pages={1083--1086},
  year={2004},
  publisher={Springer}
}

@article{rahim2002inheritance,
  title={Inheritance of resistance to sunflower downy mildew races 1, 2 and 3 in cultivated sunflower},
  author={Rahim, M and Jan, CC and Gulya, TJ},
  journal={Plant Breeding},
  volume={121},
  number={1},
  pages={57--60},
  year={2002},
  publisher={Wiley Online Library}
}

@article{miller1991inheritance,
  title={Inheritance of resistance to race 4 of downy mildew derived from interspecific crosses in sunflower},
  author={Miller, JF and Gulya, TJ},
  journal={Crop science},
  volume={31},
  number={1},
  pages={40--43},
  year={1991},
  publisher={Crop Science Society of America}
}

@article{cohen1973factors,
  title={Factors affecting infection of sunflowers by Plasmopara halstedii},
  author={Cohen, Yigal and Sackston, WE},
  journal={Canadian Journal of Botany},
  volume={51},
  number={1},
  pages={15--22},
  year={1973},
  publisher={NRC Research Press}
}

@article{heiser1978taxonomy,
  title={Taxonomy of Helianthus and origin of domesticated sunflower},
  author={Heiser, Charles B},
  journal={Sunflower science and technology, Agronomy Series},
  volume={19},
  pages={31--35},
  year={1978}
}

@book{smith2000sorghum,
  title={Sorghum: Origin, history, technology, and production},
  author={Smith, C Wayne and Frederiksen, Richard A},
  volume={2},
  year={2000},
  publisher={John Wiley \& Sons}
}

@book{doggett1988sorghum,
  title={Sorghum.},
  author={Doggett, Hugh and others},
  number={2. ed.},
  year={1988},
  publisher={Longman Scientific and Technical}
}

@article{harlan1971toward,
  title={Toward a rational classification of cultivated plants},
  author={Harlan, Jack R and de Wet, Jan MJ},
  journal={Taxon},
  pages={509--517},
  year={1971},
  publisher={JSTOR}
}

@webpage{USDAARS2007,
National Genetic Resources Program. Germplasm Resources Information Network (GRIN) [Online Database]
}

@article{garber1950cytotaxonomic,
  title={Cytotaxonomic studies in the genus Sorghum.},
  author={Garber, Edward D and others},
  journal={Univiversity of California Publications in Botany},
  volume={23},
  pages={283--362},
  year={1950}
}

@inproceedings{kuhlman2006interspecific,
  title={Interspecific sorghum breeding using S. macrospermum},
  author={Kuhlman, LC and Burson, BL and Klein, PE and Stelly, DM and Rooney, WL},
  booktitle={Proceedings of the ASA-CSSA-SSA 2006 International Meetings, Indianapolis},
  pages={12--16},
  year={2006}
}

@article{de1978systematics,
  title={Systematics and evolution of Sorghum sect. Sorghum (Gramineae)},
  author={De Wet, JMJ},
  journal={American Journal of Botany},
  pages={477--484},
  year={1978},
  publisher={JSTOR}
}

@article{yatabe2007rampant,
  title={Rampant gene exchange across a strong reproductive barrier between the annual sunflowers, Helianthus annuus and H. petiolaris},
  author={Yatabe, Yoko and Kane, Nolan C and Scotti-Saintagne, Caroline and Rieseberg, Loren H},
  journal={Genetics},
  volume={175},
  number={4},
  pages={1883--1893},
  year={2007},
  publisher={Genetics Soc America}
}

@article{schwarzbach2002likely,
  title={Likely multiple origins of a diploid hybrid sunflower species},
  author={Schwarzbach, AE and Rieseberg, LH},
  journal={Molecular Ecology},
  volume={11},
  number={9},
  pages={1703--1715},
  year={2002},
  publisher={Wiley Online Library}
}

@article{welch2002patterns,
  title={Patterns of genetic variation suggest a single, ancient origin for the diploid hybrid species Helianthus paradoxus},
  author={Welch, Mark E and Rieseberg, Loren H},
  journal={Evolution},
  volume={56},
  number={11},
  pages={2126--2137},
  year={2002},
  publisher={Wiley Online Library}
}

@article{saucă2010introgression,
  title={Introgression of drought-resistance gene (s) from Helianthus argophyllus to Helianthus annuus specie, using embryo rescue techniques.},
  author={Saucă, FLORENTINA and others},
  journal={Romanian Agricultural Research},
  number={27},
  pages={47--51},
  year={2010},
  publisher={National Agricultural Research and Development Institute}
}

@article{besnard1997specifying,
  title={Specifying the introgressed regions from H. argophyllus in cultivated sunflower (Helianthus annuus L.) to mark Phomopsis resistance genes},
  author={Besnard, G and Griveau, Y and Quillet, MC and Serieys, H and Lambert, P and Vares, D and Bervill{\'e}, A},
  journal={Theoretical and applied genetics},
  volume={94},
  number={1},
  pages={131--138},
  year={1997},
  publisher={Springer}
}

@article{shechter1975comparative,
  title={Comparative electrophoresis and isozyme analysis of seed proteins from cultivated races of sorghum},
  author={Shechter, Yaakov and De Wet, JMJ},
  journal={American Journal of Botany},
  pages={254--261},
  year={1975},
  publisher={JSTOR}
}

@article{po1982sorghum,
  title={Sorghum in the Eighties},
  author={PO, ICRISAT Patancheru},
  year={1982}
}

@article{gebrekidan1979sorghum,
  title={Sorghum Genetic Resources in Africa.},
  author={Gebrekidan, Brhane},
  journal={Ethiopian Journal of Agricultural Sciences},
  year={1979}
}

@article{reddy2006current,
  title={Current status of sorghum genetic resources at ICRISAT: their sharing and impacts},
  author={Reddy, V Gopal and Upadhyaya, HD and Gowda, CLL},
  journal={International Sorghum and Millets Newsletter},
  volume={47},
  pages={9--13},
  year={2006}
}

@article{johnson1979breeding,
  title={Breeding for arthropod resistance in sorghum},
  author={Johnson, Jerry W and Teetes, George L},
  journal={Harris, MK Biology and breeding for resistance to arthropods and pathogens in agricultural plants. Texas, USA},
  pages={168--180},
  year={1979}
}

@article{de1976cytogenetics,
  title={Cytogenetics of introgression from Saccharum into Sorghum},
  author={De Wet, JMJ and Gupta, SC and Harlan, JR and Grassl, CO},
  journal={Crop Science},
  volume={16},
  number={4},
  pages={568--572},
  year={1976},
  publisher={Crop Science Society of America}
}

@article{pickersgill1977origins,
  title={Origins and distribution of plants domesticated in the New World tropics},
  author={Pickersgill, Barbara and Heiser Jr, Charles B},
  journal={Origins of agriculture},
  year={1977}
}

@article{brush1995potato,
  title={Potato diversity in the Andean center of crop domestication},
  author={Brush, Stephen and Kesseli, Rick and Ortega, Ramiro and Cisneros, Pedro and Zimmerer, Karl and Quiros, Carlos},
  journal={Conservation Biology},
  volume={9},
  number={5},
  pages={1189--1198},
  year={1995},
  publisher={Wiley Online Library}
}

@article{rabinowitz1990high,
  title={High levels of interspecific hybridization betweenSolanum sparsipilum andS. stenotomum in experimental plots in the Andes},
  author={Rabinowitz, D and Linder, CR and Ortega, R and Begazo, D and Murguia, H and Douches, DS and Quiros, CF},
  journal={American Potato Journal},
  volume={67},
  number={2},
  pages={73--81},
  year={1990},
  publisher={Springer}
}

@article{johns1987relationships,
  title={Relationships among wild, weed, and cultivated potatoes in the Solanum x ajanhuiri complex},
  author={Johns, Timothy and Huaman, Zosimo and Ochoa, Carlos and Schmiediche, Peter E},
  journal={Systematic Botany},
  pages={541--552},
  year={1987},
  publisher={JSTOR}
}

@book{linder1987diversity,
  title={Diversity within traditional potato agriculture in the Peruvian Andes and its relevance to crop evolution},
  author={Linder, Craig Randal},
  year={1987},
  publisher={Cornell University, Aug.}
}

@article{johns1986ongoing,
  title={Ongoing evolution of the potato on the altiplano of western Bolivia},
  author={Johns, Timothy and Keen, Susan L},
  journal={Economic Botany},
  volume={40},
  number={4},
  pages={409--424},
  year={1986},
  publisher={Springer}
}

@article{hawkes1962origin,
  title={ORIGIN OF SOLANUM JUZEPCZUKII BUK AND S CURTILOBUM JUZ ET BUK},
  author={Hawkes, John Gregory},
  journal={ZEITSCHRIFT FUR PFLANZENZUCHTUNG-JOURNAL OF PLANT BREEDING},
  volume={47},
  number={1},
  pages={1},
  year={1962},
  publisher={Verlag Paul Parey SPITALERST. 12, D-2000 HAMBURG 1, GERMANY}
}

@article{schmiediche1980breeding,
  title={Breeding of the cultivated potato species Solanum x juzepczukii Buk. and Solanum x curtilobum Juz. etBuk.},
  author={Schmiediche, PE and Hawkes, JG and Ochoa, CM},
  journal={Euphytica},
  volume={29},
  number={3},
  pages={685--704},
  year={1980},
  publisher={Springer}
}

@article{brush1981dynamics,
  title={Dynamics of Andean potato agriculture},
  author={Brush, Stephen B and Carney, Heath J and Hum{\'a}n, Z{\'o}simo},
  journal={Economic Botany},
  volume={35},
  number={1},
  pages={70--88},
  year={1981},
  publisher={Springer}
}

@article{hawkes1988evolution,
  title={The evolution of cultivated potatoes and their tuber-bearing wild relatives},
  author={Hawkes, John G},
  journal={Die Kulturpflanze},
  volume={36},
  number={1},
  pages={189--208},
  year={1988},
  publisher={Springer}
}

@article{kamala2012interactions,
  title={Interactions of spotted stem borer Chilo partellus with wild relatives of sorghum},
  author={Kamala, Venkateswaran and Sharma, Hari C and Manohar Rao, Daggu and Varaprasad, Kodeboyina S and Bramel, Paula J and Chandra, Subhash},
  journal={Plant breeding},
  volume={131},
  number={4},
  pages={511--521},
  year={2012},
  publisher={Wiley Online Library}
}

@book{ellstrand2003dangerous,
  title={Dangerous liaisons?: when cultivated plants mate with their wild relatives},
  author={Ellstrand, Norman Carl},
  year={2003},
  publisher={JHU Press}
}

@article{harlan1995living,
  title={The living fields},
  author={Harlan, Jack R},
  journal={Our agricultural heritage.-271 S},
  year={1995}
}

@article{doggett1968disruptive,
  title={Disruptive selection in crop development},
  author={Doggett, H and Majisu, BN},
  journal={Heredity},
  volume={23},
  number={1},
  pages={1--23},
  year={1968}
}

@article{baker1972human,
  title={Human influences on plant evolution},
  author={Baker, Herbert G},
  journal={Economic Botany},
  volume={26},
  number={1},
  pages={32--43},
  year={1972},
  publisher={Springer}
}

@article{aldrich1992restriction,
  title={Restriction fragment variation in the nuclear and chloroplast genomes of cultivated and wild Sorghum bicolor},
  author={Aldrich, PR and Doebley, J},
  journal={Theoretical and Applied Genetics},
  volume={85},
  number={2-3},
  pages={293--302},
  year={1992},
  publisher={Springer}
}

@article{aldrich1992patterns,
  title={Patterns of allozyme variation in cultivated and wild Sorghum bicolor},
  author={Aldrich, PR and Doebley, J and Schertz, KF and Stec, A},
  journal={Theoretical and Applied Genetics},
  volume={85},
  number={4},
  pages={451--460},
  year={1992},
  publisher={Springer}
}

@article{arriola1996crop,
  title={Crop-to-weed gene flow in the genus Sorghum (Poaceae): spontaneous interspecific hybridization between johnsongrass, Sorghum halepense, and crop sorghum, S. bicolor},
  author={Arriola, Paul E and Ellstrand, Norman C},
  journal={American Journal of Botany},
  pages={1153--1159},
  year={1996},
  publisher={JSTOR}
}

@article{kwak2009structure,
  title={Structure of genetic diversity in the two major gene pools of common bean (Phaseolus vulgaris L., Fabaceae)},
  author={Kwak, Myounghai and Gepts, Paul},
  journal={Theoretical and Applied Genetics},
  volume={118},
  number={5},
  pages={979--992},
  year={2009},
  publisher={Springer}
}

@article{molina2011molecular,
  title={Molecular evidence for a single evolutionary origin of domesticated rice},
  author={Molina, Jeanmaire and Sikora, Martin and Garud, Nandita and Flowers, Jonathan M and Rubinstein, Samara and Reynolds, Andy and Huang, Pu and Jackson, Scott and Schaal, Barbara A and Bustamante, Carlos D and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={20},
  pages={8351--8356},
  year={2011},
  publisher={National Acad Sciences}
}

@article{zhao2010genomic,
  title={Genomic diversity and introgression in O. sativa reveal the impact of domestication and breeding on the rice genome},
  author={Zhao, Keyan and Wright, Mark and Kimball, Jennifer and Eizenga, Georgia and McClung, Anna and Kovach, Michael and Tyagi, Wricha and Ali, M Likat and Tung, Chih-Wei and Reynolds, Andy and others},
  journal={PloS one},
  volume={5},
  number={5},
  pages={e10780},
  year={2010}
}

@article{chang2003origin,
  title={Origin, domestication, and diversification},
  author={Chang, Te-Tzu},
  journal={Rice: origin, history, technology, and production. New Jersey USA: J. Wiley and Sons, Inc},
  pages={3--25},
  year={2003}
}

@article{khush2003classifying,
  title={Classifying rice germplasm by isozyme polymorphism and origin of cultivated rice},
  author={Khush, GS and Brar, DS and Virk, PS and Tang, SX and Malik, SS and Busto, GA and Lee, YT and McNally, R and Trinh, LN and Jiang, Y and others},
  journal={IRRI Discuss Pap},
  volume={46},
  pages={279},
  year={2003}
}

@article{glaszmann1987isozymes,
  title={Isozymes and classification of Asian rice varieties},
  author={Glaszmann, JC},
  journal={Theoretical and Applied Genetics},
  volume={74},
  number={1},
  pages={21--30},
  year={1987},
  publisher={Springer}
}

@article{zohary1959hordeum,
  title={Is Hordeum a griocrithon the Ancestor of Six-Rowed Cultivated Barley?},
  author={Zohary, Daniel},
  journal={Evolution},
  volume={13},
  number={2},
  pages={279--280},
  year={1959},
  publisher={JSTOR}
}

@article{staudt1961origin,
  title={The origin of cultivated barleys: a discussion},
  author={Staudt, G},
  journal={Economic Botany},
  volume={15},
  number={3},
  pages={205--212},
  year={1961},
  publisher={Springer}
}

@article{salamini2002genetics,
  title={Genetics and geography of wild cereal domestication in the Near East},
  author={Salamini, Francesco and {\"O}zkan, Hakan and Brandolini, Andrea and Sch{\"a}fer-Pregl, Ralf and Martin, William},
  journal={Nature Reviews Genetics},
  volume={3},
  number={6},
  pages={429--441},
  year={2002},
  publisher={Nature Publishing Group}
}

@article{civavn2013reticulated,
  title={Reticulated origin of domesticated emmer wheat supports a dynamic model for the emergence of agriculture in the fertile crescent},
  author={Civ{\'a}{\v{n}}, Peter and Ivani{\v{c}}ov{\'a}, Zuzana and Brown, Terence A},
  year={2013}
}

@book{hancock2012plant,
  title={Plant evolution and the origin of crop species},
  author={Hancock, James F},
  year={2012},
  publisher={CABI}
}

@article{hillman2001new,
  title={New evidence of Lateglacial cereal cultivation at Abu Hureyra on the Euphrates},
  author={Hillman, Gordon and Hedges, Robert and Moore, Andrew and Colledge, Susan and Pettitt, Paul},
  journal={The Holocene},
  volume={11},
  number={4},
  pages={383--393},
  year={2001},
  publisher={Sage Publications}
}

@article{khush1961cytogenetic,
  title={Cytogenetic and evolutionary studies in Secale. I. Some new data on the ancestry of S. cereale},
  author={Khush, Gurdev S and Stebbins, G Ledyard},
  journal={American Journal of Botany},
  pages={723--730},
  year={1961},
  publisher={JSTOR}
}

@article{kane2007selective,
  title={Selective sweeps reveal candidate genes for adaptation to drought and salt tolerance in common sunflower, Helianthus annuus},
  author={Kane, Nolan C and Rieseberg, Loren H},
  journal={Genetics},
  volume={175},
  number={4},
  pages={1823--1834},
  year={2007},
  publisher={Genetics Soc America}
}

@article{rieseberg1990helianthus,
  title={Helianthus annuus ssp. texanus has chloroplast DNA and nuclear ribosomal RNA genes of Helianthus debilis ssp. cucumerifolius},
  author={Rieseberg, Loren H and Beckstrom-Sternberg, Stephen and Doan, Khiem},
  journal={Proceedings of the National Academy of Sciences},
  volume={87},
  number={2},
  pages={593--597},
  year={1990},
  publisher={National Acad Sciences}
}

@article{stutz1972origin,
  title={On the origin of cultivated rye},
  author={Stutz, Howard C},
  journal={American Journal of Botany},
  pages={59--70},
  year={1972},
  publisher={JSTOR}
}

@article{fuller2010consilience,
  title={Consilience of genetics and archaeobotany in the entangled history of rice},
  author={Fuller, Dorian Q and Sato, Yo-Ichiro and Castillo, Cristina and Qin, Ling and Weisskopf, Alison R and Kingwell-Banham, Eleanor J and Song, Jixiang and Ahn, Sung-Mo and Van Etten, Jacob},
  journal={Archaeological and Anthropological Sciences},
  volume={2},
  number={2},
  pages={115--131},
  year={2010},
  publisher={Springer}
}

@article{baker1972migrations,
  title={Migrations of weeds},
  author={Baker, Herbert G},
  journal={Valentine, D, H ed (s). Taxonomy phytogeography and evolution},
  pages={327--347},
  year={1972}
}

@inproceedings{vavilov1928geographical,
  title={Geographical centers of our cultivated plants},
  author={Vavilov, NI},
  booktitle={Proceedings of V international congress on genetics, New York, USA},
  pages={342--369},
  year={1928}
}

@article{quiros1992increase,
  title={Increase of potato genetic resources in their center of diversity: the role of natural outcrossing and selection by the Andean farmer},
  author={Quiros, CF and Ortega, R and Van Raamsdonk, L and Herrera-Montoya, M and Cisneros, P and Schmidt, E and Brush, SB},
  journal={Genetic Resources and Crop Evolution},
  volume={39},
  number={2},
  pages={107--113},
  year={1992},
  publisher={Springer}
}

@article{zimmerer1998ecogeography,
  title={The ecogeography of Andean potatoes},
  author={Zimmerer, Karl S},
  journal={BioScience},
  pages={445--454},
  year={1998},
  publisher={JSTOR}
}

@article{zimmerer1991geographical,
  title={Geographical approaches to crop conservation: the partitioning of genetic diversity in Andean potatoes},
  author={Zimmerer, Karl S and Douches, David S},
  journal={Economic Botany},
  volume={45},
  number={2},
  pages={176--189},
  year={1991},
  publisher={Springer}
}

@article{hamrick1983distribution,
  title={The distribution of genetic variation within and among natural plant populations.},
  author={Hamrick, James L},
  journal={BIOL. CONSERV. SER.},
  pages={335--348},
  year={1983}
}

@article{hammer1987collection,
  title={Collection of plant genetic resources in South Italy, 1986},
  author={Hammer, Karl and Perrino, Pietro and Pignone, Domenico},
  journal={Die Kulturpflanze},
  volume={35},
  number={2},
  pages={389--399},
  year={1987},
  publisher={Springer}
}

@article{murphy1982origin,
  title={The origin of six-rowed ‘wild’barley from the western Himalaya},
  author={Murphy, PJ and Witcombe, JR and Shewry, PR and Miflin, BJ},
  journal={Euphytica},
  volume={31},
  number={1},
  pages={183--192},
  year={1982},
  publisher={Springer}
}

@article{huaman1980solanum,
  title={Solanum ajanhuiri: an important diploid potato cultivated in the Andean altiplano},
  author={Huam{\'a}n, Z and Hawkes, JG and Rowe, PR},
  journal={Economic Botany},
  volume={34},
  number={4},
  pages={335--343},
  year={1980},
  publisher={Springer}
}

@article{carney2000evolutionary,
  title={Evolutionary changes over the fifty-year history of a hybrid population of sunflowers (Helianthus)},
  author={Carney, Shanna E and Gardner, Keith A and Rieseberg, Loren H},
  journal={Evolution},
  volume={54},
  number={2},
  pages={462--474},
  year={2000},
  publisher={BioOne}
}

@article{rieseberg1988molecular,
  title={A molecular reexamination of introgression between Helianthus annuus and H bolanderi (Compositae)},
  author={Rieseberg, Loren H and Soltis, Douglas E and Palmer, Jeffrey D},
  journal={Evolution},
  pages={227--238},
  year={1988},
  publisher={JSTOR}
}

@article{gerechter1971transfer,
  title={Transfer of stem rust seedling resistance from wild diploid einkorn to tetraploid durum wheat by means of a triploid hybrid bridge},
  author={Gerechter-Amitai, ZK and Wahl, I and Vardi, Aliza and Zohary, D},
  journal={Euphytica},
  volume={20},
  number={2},
  pages={281--285},
  year={1971},
  publisher={Springer}
}

@article{blanco2008molecular,
  title={Molecular mapping of the novel powdery mildew resistance gene Pm36 introgressed from Triticum turgidum var. dicoccoides in durum wheat},
  author={Blanco, Antonio and Gadaleta, A and Cenci, A and Carluccio, AV and Abdelbacki, AMM and Simeone, R},
  journal={Theoretical and Applied Genetics},
  volume={117},
  number={1},
  pages={135--142},
  year={2008},
  publisher={Springer}
}

@article{kane2008genetics,
  title={Genetics and evolution of weedy Helianthus annuus populations: adaptation of an agricultural weed},
  author={Kane, Nolan C and Rieseberg, Loren H},
  journal={Molecular Ecology},
  volume={17},
  number={1},
  year={2008},
  publisher={Wiley Online Library}
}

@article{maron2004rapid,
  title={Rapid evolution of an invasive plant},
  author={Maron, John L and Vil{\`a}, Montserrat and Bommarco, Riccardo and Elmendorf, Sarah and Beardsley, Paul},
  journal={Ecological Monographs},
  volume={74},
  number={2},
  pages={261--280},
  year={2004},
  publisher={Eco Soc America}
}

@article{ozkan2002aflp,
  title={AFLP analysis of a collection of tetraploid wheats indicates the origin of emmer and hard wheat domestication in southeast Turkey},
  author={{\"O}zkan, Hakan and Brandolini, Andrea and Sch{\"a}fer-Pregl, Ralf and Salamini, Francesco},
  journal={Molecular Biology and Evolution},
  volume={19},
  number={10},
  pages={1797--1801},
  year={2002},
  publisher={SMBE}
}

@article{luo2007structure,
  title={The structure of wild and domesticated emmer wheat populations, gene flow between them, and the site of emmer domestication},
  author={Luo, M-C and Yang, Z-L and You, Frank M and Kawahara, Taihachi and Waines, J Giles and Dvorak, J},
  journal={Theoretical and Applied Genetics},
  volume={114},
  number={6},
  pages={947--959},
  year={2007},
  publisher={Springer}
}

@article{dvorak2006molecular,
  title={Molecular characterization of a diagnostic DNA marker for domesticated tetraploid wheat provides evidence for gene flow from wild tetraploid wheat to hexaploid wheat},
  author={Dvorak, Jan and Akhunov, Eduard D and Akhunov, Alina R and Deal, Karin R and Luo, Ming-Cheng},
  journal={Molecular biology and evolution},
  volume={23},
  number={7},
  pages={1386--1396},
  year={2006},
  publisher={SMBE}
}

@article{barton2001role,
  title={The role of hybridization in evolution},
  author={Barton, NH},
  journal={Molecular Ecology},
  volume={10},
  number={3},
  pages={551--568},
  year={2001},
  publisher={Wiley Online Library}
}

@article{stebbins1969significance,
  title={The significance of hybridization for plant taxonomy and evolution},
  author={Stebbins, G Ledyard},
  journal={Taxon},
  pages={26--35},
  year={1969},
  publisher={JSTOR}
}

@article{grun1990evolution,
  title={The evolution of cultivated potatoes},
  author={Grun, Paul},
  journal={Economic Botany},
  volume={44},
  number={3},
  pages={39--55},
  year={1990},
  publisher={Springer}
}

@article{ricepedia,
    url={http://ricepedia.org/rice-as-a-plant/rice-species/cultivated-rice-species},
}

@article{lentz2008sunflower,
  title={Sunflower (Helianthus annuus L.) as a pre-Columbian domesticate in Mexico},
  author={Lentz, David L and Pohl, Mary DeLand and Alvarado, Jos{\'e} Luis and Tarighat, Somayeh and Bye, Robert},
  journal={Proceedings of the National Academy of Sciences},
  volume={105},
  number={17},
  pages={6232--6237},
  year={2008},
  publisher={National Acad Sciences}
}

@article{harter2004origin,
  title={Origin of extant domesticated sunflowers in eastern North America},
  author={Harter, Abigail V and Gardner, Keith A and Falush, Daniel and Lentz, David L and Bye, Robert A and Rieseberg, Loren H},
  journal={Nature},
  volume={430},
  number={6996},
  pages={201--205},
  year={2004},
  publisher={Nature Publishing Group}
}

@article{mcfadden1946origin,
  title={The origin of Triticum spelta and its free-threshing hexaploid relatives},
  author={McFadden, ES and Sears, ER},
  journal={Journal of Heredity},
  volume={37},
  number={4},
  pages={107--116},
  year={1946},
  publisher={Am Genetic Assoc}
}

@article{marcussen2014ancient,
  title={Ancient hybridizations among the ancestral genomes of bread wheat},
  author={Marcussen, Thomas and Sandve, Simen R and Heier, Lise and Spannagl, Manuel and Pfeifer, Matthias and Jakobsen, Kjetill S and Wulff, Brande BH and Steuernagel, Burkhard and Mayer, Klaus FX and Olsen, Odd-Arne and others},
  journal={Science},
  volume={345},
  number={6194},
  pages={1250092},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{petersen2006phylogenetic,
  title={Phylogenetic relationships of Triticum and Aegilops and evidence for the origin of the A, B, and D genomes of common wheat (Triticum aestivum)},
  author={Petersen, Gitte and Seberg, Ole and Yde, Merete and Berthelsen, Kasper},
  journal={Molecular phylogenetics and evolution},
  volume={39},
  number={1},
  pages={70--82},
  year={2006},
  publisher={Elsevier}
}

@article{haberer2015barley,
  title={Barley: From Brittle to Stable Harvest},
  author={Haberer, Georg and Mayer, Klaus FX},
  journal={Cell},
  volume={162},
  number={3},
  pages={469--471},
  year={2015},
  publisher={Elsevier}
}

@article{giles2006gludy,
  title={GluDy allele variations in Aegilops tauschii and Triticum aestivum: implications for the origins of hexaploid wheats},
  author={Giles, Rachel J and Brown, Terence A},
  journal={Theoretical and applied genetics},
  volume={112},
  number={8},
  pages={1563--1572},
  year={2006},
  publisher={Springer}
}

@article{ting1933wild,
  title={Wild rice of Kwangtung and new variety bred from the hybrids of wild rice with cultivated rice},
  author={Ting, YING},
  journal={Coll. Agric. Sun Yatsen Univ., Agron. Bull},
  volume={3},
  pages={1--22},
  year={1933}
}

@webpage{CWR,
    url={http://www.cwrdiversity.org/distribution-map/}
}

@article{londo2006phylogeography,
  title={Phylogeography of Asian wild rice, Oryza rufipogon, reveals multiple independent domestications of cultivated rice, Oryza sativa},
  author={Londo, Jason P and Chiang, Yu-Chung and Hung, Kuo-Hsiang and Chiang, Tzen-Yuh and Schaal, Barbara A},
  journal={Proceedings of the National Academy of Sciences},
  volume={103},
  number={25},
  pages={9578--9583},
  year={2006},
  publisher={National Acad Sciences}
}

@article{zohary1975beginnings,
  title={Beginnings of fruit growing in the Old World},
  author={Zohary, Daniel and Spiegel-Roy, Pinhas},
  journal={Science},
  volume={187},
  number={4174},
  pages={319--327},
  year={1975},
  publisher={American Association for the Advancement of Science}
}

@article{austin1988taxonomy,
  title={The taxonomy, evolution and genetic diversity of sweet potatoes and related wild species},
  author={Austin, Daniel F},
  journal={Exploration, maintenance, and utilization of sweetpotato genetic resources. International Potato Center, Lima, Peru},
  pages={27--60},
  year={1988}
}

@phdthesis{akinbo2008introgression,
  title={Introgression of high protein and pest resistance genes from inter-specific hybrids of Manihot esculenta ssp flabellifolia into cassava (Manihot esculenta Crantz)},
  author={Akinbo, Olalekan Abiodun},
  year={2008},
  school={University of the Free State}
}

@article{tovar2015diversity,
  title={Diversity and genetic structure of cassava landraces and their wild relatives (Manihot spp.) in Colombia revealed by simple sequence repeats},
  author={Tovar, E and Bocanegra, JL and Villafa{\~n}e, C and Fory, L and Vel{\'a}squez, A and Gallego, G and Moreno, R},
  journal={Plant Genetic Resources},
  pages={1--11},
  year={2015},
  publisher={Cambridge Univ Press}
}

@article{roullier2013disentangling,
  title={Disentangling the origins of cultivated sweet potato (Ipomoea batatas (L.) Lam.)},
  author={Roullier, Caroline and Duputi{\'e}, Anne and Wennekes, Paul and Benoit, Laure and Bringas, V{\'\i}ctor Manuel Fern{\'a}ndez and Rossel, Genoveva and Tay, David and McKey, Doyle and Lebot, Vincent},
  journal={PLoS One},
  volume={8},
  number={5},
  pages={e62707},
  year={2013},
  publisher={Public Library of Science}
}

@article{roa1997aflp,
  title={AFLP analysis of relationships among cassava and other Manihot species},
  author={Roa, Ana Carolina and Maya, Mar{\'\i}a Mercedes and Duque, MC and Tohme, J and Allem, Antonio Costa and Bonierbale, Merideth W},
  journal={Theoretical and applied Genetics},
  volume={95},
  number={5-6},
  pages={741--750},
  year={1997},
  publisher={Springer}
}

@article{olsen1999evidence,
  title={Evidence on the origin of cassava: phylogeography of Manihot esculenta},
  author={Olsen, Kenneth M and Schaal, Barbara A},
  journal={Proceedings of the National Academy of Sciences},
  volume={96},
  number={10},
  pages={5586--5591},
  year={1999},
  publisher={National Acad Sciences}
}

@article{daniels1987taxonomy,
  title={Taxonomy and evolution},
  author={Daniels, J and Roach, Brian T},
  journal={Sugarcane improvement through breeding. Elsevier, Amsterdam},
  volume={7},
  pages={7--84},
  year={1987}
}

@article{jackson2005breeding,
  title={Breeding for improved sugar content in sugarcane},
  author={Jackson, Phillip A},
  journal={Field Crops Research},
  volume={92},
  number={2},
  pages={277--290},
  year={2005},
  publisher={Elsevier}
}

@article{brandes1958origin,
  title={Origin, classification and characteristics},
  author={Brandes, EW},
  journal={Sugarcane},
  pages={1--35},
  year={1958}
}

@article{irvine1999saccharum,
  title={Saccharum species as horticultural classes},
  author={Irvine, JE},
  journal={Theoretical and Applied Genetics},
  volume={98},
  number={2},
  pages={186--194},
  year={1999},
  publisher={Springer}
}

@inproceedings{roach1989origin,
  title={Origin and improvement of the genetic base of sugarcane},
  author={Roach, BT},
  booktitle={Proc Aust Soc Sugar Cane Technol},
  volume={11},
  pages={34--47},
  year={1989}
}

@webpage{keyroyalbotanicalgardens},
  url={http://www.kew.org/science-conservation/plants-fungi/saccharum-officinarum-sugar-cane},
}

@article{artschwager1958sugarcane,
  title={Sugarcane (Saccharum officinarum L.).},
  author={Artschwager, Ernst and Brandes, Elmer Walker and others},
  journal={Sugarcane (Saccharum officinarum L.).},
  year={1958},
  publisher={USDA}
}

@article{daniels1987taxonomy,
  title={Taxonomy and evolution},
  author={Daniels, J and Roach, Brian T},
  journal={Sugarcane improvement through breeding. Elsevier, Amsterdam},
  volume={7},
  pages={7--84},
  year={1987}
}

@article{roach1995sugar,
  title={Sugar canes--Saccharum (Gramineae-Andropogonae)},
  author={Roach, BT},
  year={1995},
  publisher={in: in: J. Smartt and NW Simmonds (Eds.), 160--166}
}

@article{sobral1994phylogenetic,
  title={Phylogenetic analysis of chloroplast restriction enzyme site mutations in the Saccharinae Griseb. subtribe of the Andropogoneae Dumort. tribe},
  author={Sobral, BWS and Braga, DPV and LaHood, ES and Keim, P},
  journal={Theoretical and Applied Genetics},
  volume={87},
  number={7},
  pages={843--853},
  year={1994},
  publisher={Springer}
}

@techreport{stevenson1965genetics,
  title={Genetics and breeding of sugar cane},
  author={Stevenson, Griffith Cross and others},
  year={1965},
  institution={Cambridge Univ Press}
}

@article{ellstrand1999gene,
  title={Gene flow and introgression from domesticated plants into their wild relatives},
  author={Ellstrand, Norman C and Prentice, Honor C and Hancock, James F},
  journal={Annual review of Ecology and Systematics},
  pages={539--563},
  year={1999},
  publisher={JSTOR}
}

#Li had these in a separate reference.bib file on Github:

@article{andrew2012sunflower,
  title={Adaptation with gene flow across the landscape in a dune sunflower},
  author={Andrew, Rose L and Ostevik, Katherine L and Ebert, Daniel P and Rieseberg, Loren H},
  journal={Molecular ecology},
  volume={21},
  number={9},
  pages={2078--2091},
  year={2012},
  publisher={Wiley Online Library}
}

@article{feder2012,
  title={The genomics of speciation-with-gene-flow},
  author={Feder, Jeffrey L and Egan, Scott P and Nosil, Patrik},
  journal={Trends in Genetics},
  volume={28},
  number={7},
  pages={342--350},
  year={2012},
  publisher={Elsevier}
}

@article{racimo2015,
  title={Evidence for archaic adaptive introgression in humans},
  author={Racimo, Fernando and Sankararaman, Sriram and Nielsen, Rasmus and Huerta-S{\'a}nchez, Emilia},
  journal={Nature Reviews Genetics},
  volume={16},
  number={6},
  pages={359--371},
  year={2015},
  publisher={Nature Publishing Group}
}

@article{Takuno2015,
author = {Takuno, Shohei and Ralph, Peter and Swarts, Kelly and Elshire, Rob J. and Glaubitz, Jeffrey C. and Buckler, Edward S. and Hufford, Matthew B. and Ross-Ibarra, Jeffrey}, 
title = {Independent Molecular Basis of Convergent Highland Adaptation in Maize},
year = {2015}, 
doi = {10.1534/genetics.115.178327}, 
URL = {http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.abstract}, 
eprint = {http://www.genetics.org/content/early/2015/06/15/genetics.115.178327.full.pdf+html}, 
journal = {Genetics} 
}

@article{hufford2013,
  title={The genomic signature of crop-wild introgression in maize},
  author={Hufford, Matthew B and Lubinksy, Pesach and Pyh{\"a}j{\"a}rvi, Tanja and Devengenzo, Michael T and Ellstrand, Norman C and Ross-Ibarra, Jeffrey},
  year={2013},
  doi={10.1371/journal.pgen.1003477},
  journal={Plos genetics} 
}

@article{vanHeerwaarden2011,
  title={Genetic signals of origin, spread, and introgression in a large sample of maize landraces},
  author={van Heerwaarden, Joost and Doebley, John and Briggs, William H and Glaubitz, Jeffrey C and Goodman, Major M and Gonzalez, Jose de Jesus Sanchez and Ross-Ibarra, Jeffrey},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={3},
  pages={1088--1092},
  year={2011},
  publisher={National Acad Sciences}
}

@article{matsuoka2002,
  title={A single domestication for maize shown by multilocus microsatellite genotyping},
  author={Matsuoka, Yoshihiro and Vigouroux, Yves and Goodman, Major M and Sanchez, Jesus and Buckler, Edward and Doebley, John},
  journal={Proceedings of the National Academy of Sciences},
  volume={99},
  number={9},
  pages={6080--6084},
  year={2002},
  publisher={National Acad Sciences}
}

@article{wilkes1977,
  title={Hybridization of maize and teosinte, in Mexico and Guatemala and the improvement of maize},
  author={Wilkes, HG},
  journal={Economic Botany},
  pages={254--293},
  year={1977},
  publisher={JSTOR}
}

@article{lauter2004,
  title={The inheritance and evolution of leaf pigmentation and pubescence in teosinte},
  author={Lauter, Nick and Gustus, Charles and Westerbergh, Anna and Doebley, John},
  journal={Genetics},
  volume={167},
  number={4},
  pages={1949--1959},
  year={2004},
  publisher={Genetics Soc America}
}

@article{doebley1984,
  title={Maize introgression into teosinte-a reappraisal},
  author={Doebley, John F},
  journal={Annals of the Missouri Botanical Garden},
  pages={1100--1113},
  year={1984},
  publisher={JSTOR}
}

@article{doebley1987,
  title={Patterns of isozyme variation between maize and Mexican annual teosinte},
  author={Doebley, John and Goodman, Major M and Stuber, Charles W},
  journal={Economic Botany},
  volume={41},
  number={2},
  pages={234--246},
  year={1987},
  publisher={Springer}
}

@article{warburton2011,
  title={Gene flow among different teosinte taxa and into the domesticated maize gene pool},
  author={Warburton, Marilyn L and Wilkes, Garrison and Taba, S and Charcosset, Alain and Mir, Celine and Dumas, Fabrice and Madur, Delphine and Dreisigacker, Susanne and Bedoya, Claudia and Prasanna, BM and others},
  journal={Genetic resources and crop evolution},
  volume={58},
  number={8},
  pages={1243--1261},
  year={2011},
  publisher={Springer}
}

@article{fukunaga2005,
  title={Genetic diversity and population structure of teosinte},
  author={Fukunaga, Kenji and Hill, Jason and Vigouroux, Yves and Matsuoka, Yoshihiro and Sanchez, Jesus and Liu, Kejun and Buckler, Edward S and Doebley, John},
  journal={Genetics},
  volume={169},
  number={4},
  pages={2241--2254},
  year={2005},
  publisher={Genetics Soc America}
}

@article{martin2015,
  title={Evaluating the use of ABBA--BABA statistics to locate introgressed loci},
  author={Martin, Simon H and Davey, John W and Jiggins, Chris D},
  journal={Molecular biology and evolution},
  volume={32},
  number={1},
  pages={244--257},
  year={2015},
  publisher={SMBE}
}

#The following were in the TiG prospectus bib file.

@Article{Hufford2013,
author = {Hufford, MB and Lubinksy, P and Pyh\"{a}j\"{a}rvi, T and Devengenzo, MT and Ellstrand, NC and Ross-Ibarra, J}, 
title = {The genomic signature of crop-wild introgression in maize}, 
journal = {PLoS Genetics}, 
volume = {9}, 
number = {5}, 
pages = {e1003477}, 
year = {2013}, 
abstract = {The evolutionary significance of hybridization and subsequent introgression has long been appreciated, but evaluation of the genome-wide effects of these phenomena has only recently become possible. Crop-wild study systems represent ideal opportunities to examine evolution through hybridization. For example, maize and the conspecific wild teosinte Zea mays ssp. mexicana (hereafter, mexicana) are known to hybridize in the fields of highland Mexico. Despite widespread evidence of gene flow, maize and mexicana maintain distinct morphologies and have done so in sympatry for thousands of years. Neither the genomic extent nor the evolutionary importance of introgression between these taxa is understood. In this study we assessed patterns of genome-wide introgression based on 39,029 single nucleotide polymorphisms genotyped in 189 individuals from nine sympatric maize-mexicana populations and reference allopatric populations. While portions of the maize and mexicana genomes appeared resistant to introgression (notably near known cross-incompatibility and domestication loci), we detected widespread evidence for introgression in both directions of gene flow. Through further characterization of these genomic regions and preliminary growth chamber experiments, we found evidence suggestive of the incorporation of adaptive mexicana alleles into maize during its expansion to the highlands of central Mexico. In contrast, very little evidence was found for adaptive introgression from maize to mexicana. The methods we have applied here can be replicated widely, and such analyses have the potential to greatly inform our understanding of evolution through introgressive hybridization. Crop species, due to their exceptional genomic resources and frequent histories of spread into sympatry with relatives, should be particularly influential in these studies.}, 
location = {Department of Plant Sciences, University of California Davis, Davis, California, United States of America.}, 
keywords = {}}

@article{Racimo2015,
	Author = {Racimo, Fernando and Sankararaman, Sriram and Nielsen, Rasmus and Huerta-Sanchez, Emilia},
	Journal = {Nat Rev Genet},
	M3 = {Review},
	Number = {6},
	Pages = {359--371},
	Title = {Evidence for archaic adaptive introgression in humans},
	Ty = {JOUR},
	Volume = {16},
	Year = {2015}
	}

@article {Baute2015,
author = {Baute, Gregory J. and Kane, Nolan C. and Grassa, Christopher J. and Lai, Zhao and Rieseberg, Loren H.},
title = {Genome scans reveal candidate domestication and improvement genes in cultivated sunflower, as well as post-domestication introgression with wild relatives},
journal = {New Phytologist},
volume = {206},
number = {2},
pages = {830--838},
keywords = {breeding, domestication, genomics, introgression, sunflower (Helianthus spp.)},
year = {2015},
}

@article{Bredeson2016,
	Author = {Bredeson, Jessen V and Lyons, Jessica B and Prochnik, Simon E and Wu, G Albert and Ha, Cindy M and Edsinger-Gonzales, Eric and Grimwood, Jane and Schmutz, Jeremy and Rabbi, Ismail Y and Egesi, Chiedozie and Nauluvula, Poasa and Lebot, Vincent and Ndunguru, Joseph and Mkamilo, Geoffrey and Bart, Rebecca S and Setter, Tim L and Gleadow, Roslyn M and Kulakow, Peter and Ferguson, Morag E and Rounsley, Steve and Rokhsar, Daniel S},
	Journal = {Nat Biotech},
	Number = {5},
	Pages = {562--570},
	Title = {Sequencing wild and cultivated cassava and related species reveals extensive interspecific hybridization and genetic diversity},
	Volume = {34},
	Year = {2016},
	}

@article{Heliconius2012,
	Author = {The Heliconius Genome Consortium},
	Journal = {Nature},
	Number = {7405},
	Pages = {94--98},
	Title = {Butterfly genome reveals promiscuous exchange of mimicry adaptations among species},
	Volume = {487},
	Year = {2012},
	}

@article{Norris2015,
author = {Norris, Laura C. and Main, Bradley J. and Lee, Yoosook and Collier, Travis C. and Fofana, Abdrahamane and Cornel, Anthony J. and Lanzaro, Gregory C.}, 
title = {Adaptive introgression in an African malaria mosquito coincident with the increased usage of insecticide-treated bed nets},
volume = {112}, 
number = {3}, 
pages = {815--820}, 
year = {2015}, 
journal = {Proceedings of the National Academy of Sciences} 
}

@Article{Poets2015,
author="Poets, Ana M.
and Fang, Zhou
and Clegg, Michael T.
and Morrell, Peter L.",
title="Barley landraces are characterized by geographically heterogeneous genomic origins",
journal="Genome Biology",
year="2015",
volume="16",
number="1",
pages="1--11",
}

@article{Huang2012,
	Author = {Huang, Xuehui and Kurata, Nori and Wei, Xinghua and Wang, Zi-Xuan and Wang, Ahong and Zhao, Qiang and Zhao, Yan and Liu, Kunyan and Lu, Hengyun and Li, Wenjun and Guo, Yunli and Lu, Yiqi and Zhou, Congcong and Fan, Danlin and Weng, Qijun and Zhu, Chuanrang and Huang, Tao and Zhang, Lei and Wang, Yongchun and Feng, Lei and Furuumi, Hiroyasu and Kubo, Takahiko and Miyabayashi, Toshie and Yuan, Xiaoping and Xu, Qun and Dong, Guojun and Zhan, Qilin and Li, Canyang and Fujiyama, Asao and Toyoda, Atsushi and Lu, Tingting and Feng, Qi and Qian, Qian and Li, Jiayang and Han, Bin},
	Journal = {Nature},
	Pages = {497--501},
	Title = {A map of rice genome variation reveals the origin of cultivated rice},
	Volume = {490},
	Year = {2012},
	}

@article {Diez2015,
author = {Diez, Concepcion M. and Trujillo, Isabel and Martinez-Urdiroz, Nieves and Barranco, Diego and Rallo, Luis and Marfil, Pedro and Gaut, Brandon S.},
title = {Olive domestication and diversification in the Mediterranean Basin},
journal = {New Phytologist},
volume = {206},
number = {1},
pages = {436--447},
keywords = {center of origin, demographic history, fruit crops, genetic admixture, olive (Olea europaea ssp. europaea), simple-sequence repeats (SSR)},
year = {2015},
}

@article {Stankowski2015,
	author = {Stankowski, Sean and Streisfeld, Matthew A.},
	title = {Introgressive hybridization facilitates adaptive divergence in a recent radiation of monkeyflowers},
	volume = {282},
	number = {1814},
	year = {2015},
	journal = {Proceedings of the Royal Society of London B: Biological Sciences}
}

@article {Currat2008,
author = {Currat, Mathias and Ruedi, Manuel and Petit, Rémy J. and Excoffier, Laurent},
title = {The hidden side of invasions: massive introgression by local genes},
journal = {Evolution},
volume = {62},
number = {8},
pages = {1908--1920},
year = {2008}
}

@article{Martin2015,
author = {Martin, Simon H. and Davey, John W. and Jiggins, Chris D.}, 
title = {Evaluating the Use of ABBA–BABA Statistics to Locate Introgressed Loci},
volume = {32}, 
number = {1}, 
pages = {244-257}, 
year = {2015}, 
journal = {Molecular Biology and Evolution} 
}

@article{Price2009,
    author = {Price, Alkes L. AND Tandon, Arti AND Patterson, Nick AND Barnes, Kathleen C. AND Rafaels, Nicholas AND Ruczinski, Ingo AND Beaty, Terri H. AND Mathias, Rasika AND Reich, David AND Myers, Simon},
    journal = {PLoS Genet},
    title = {Sensitive Detection of Chromosomal Segments of Distinct Ancestry in Admixed Populations},
    year = {2009},
    volume = {5},
    pages = {1-18},
    number = {6}
}

@article{Lawson2012,
    author = {Lawson, Daniel John AND Hellenthal, Garrett AND Myers, Simon AND Falush, Daniel},
    journal = {PLoS Genet},
    title = {Inference of Population Structure using Dense Haplotype Data},
    year = {2012},
    volume = {8},
    pages = {1-16},
    number = {1}
}

@article{pease2015,
  title={Detection and Polarization of Introgression in a Five-taxon Phylogeny},
  author={Pease, James B and Hahn, Matthew W},
  journal={Systematic biology},
  volume={64},
  number={4},
  pages={651--662},
  year={2015},
  publisher={Oxford University Press}
}

@article{rosenzweig2016,
  title={Powerful methods for detecting introgressed regions from population genomic data},
  author={Rosenzweig, Benjamin K and Pease, James B and Besansky, Nora J and Hahn, Matthew W},
  journal={Molecular ecology},
  year={2016},
  publisher={Wiley Online Library}
}

@article{geneva2015,
  title={A new method to scan genomes for introgression in a secondary contact model},
  author={Geneva, Anthony J and Muirhead, Christina A and Kingan, Sarah B and Garrigan, Daniel},
  journal={PloS one},
  volume={10},
  number={4},
  pages={e0118621},
  year={2015},
  publisher={Public Library of Science}
}

%% Li added references
@article{schraiber2015,
  title={Methods and models for unravelling human evolutionary history},
  author={Schraiber, Joshua G and Akey, Joshua M},
  journal={Nature Reviews Genetics},
  year={2015},
  publisher={Nature Publishing Group}
}

%fd:

@article{zhang2016genome,
  title={Genome-wide introgression among distantly related Heliconius butterfly species},
  author={Zhang, Wei and Dasmahapatra, Kanchon K and Mallet, James and Moreira, Gilson RP and Kronforst, Marcus R},
  journal={Genome biology},
  volume={17},
  number={1},
  pages={1},
  year={2016},
  publisher={BioMed Central}
}

@article{malinsky2015,
  title={Genomic islands of speciation separate cichlid ecomorphs in an East African crater lake},
  author={Malinsky, Milan and Challis, Richard J and Tyers, Alexandra M and Schiffels, Stephan and Terai, Yohey and Ngatunga, Benjamin P and Miska, Eric A and Durbin, Richard and Genner, Martin J and Turner, George F},
  journal={Science},
  volume={350},
  number={6267},
  pages={1493--1498},
  year={2015},
  publisher={American Association for the Advancement of Science}
}

%Hapmix
@article{price2009,
  title={Sensitive detection of chromosomal segments of distinct ancestry in admixed populations},
  author={Price, Alkes L and Tandon, Arti and Patterson, Nick and Barnes, Kathleen C and Rafaels, Nicholas and Ruczinski, Ingo and Beaty, Terri H and Mathias, Rasika and Reich, David and Myers, Simon},
  journal={PLoS Genet},
  volume={5},
  number={6},
  pages={e1000519},
  year={2009},
  publisher={Public Library of Science}
}

@article{suarez2016,
  title={Genomic and functional approaches reveal a case of adaptive introgression from Populus balsamifera (balsam poplar) in P. trichocarpa (black cottonwood)},
  author={Suarez-Gonzalez, Adriana and Hefer, Charles A and Christe, Camille and Corea, Oliver and Lexer, Christian and Cronk, Quentin CB and Douglas, Carl J},
  journal={Molecular ecology},
  year={2016},
  publisher={Wiley Online Library}
}

%PCAdmix
@article{brisbin2012pcadmix,
  title={PCAdmix: principal components-based assignment of ancestry along each chromosome in individuals with admixed ancestry from two or more populations},
  author={Brisbin, Abra and Bryc, Katarzyna and Byrnes, Jake and Zakharia, Fouad and Omberg, Larsson and Degenhardt, Jeremiah and Reynolds, Andrew and Ostrer, Harry and Mezey, Jason G and Bustamante, Carlos D},
  journal={Human biology},
  volume={84},
  number={4},
  pages={343},
  year={2012},
  publisher={NIH Public Access}
}

@article{moreno2014genetics,
  title={The genetics of Mexico recapitulates Native American substructure and affects biomedical traits},
  author={Moreno-Estrada, Andr{\'e}s and Gignoux, Christopher R and Fern{\'a}ndez-L{\'o}pez, Juan Carlos and Zakharia, Fouad and Sikora, Martin and Contreras, Alejandra V and Acu{\~n}a-Alonzo, Victor and Sandoval, Karla and Eng, Celeste and Romero-Hidalgo, Sandra and others},
  journal={Science},
  volume={344},
  number={6189},
  pages={1280--1285},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{pugach2016complex,
  title={The complex admixture history and recent southern origins of Siberian populations},
  author={Pugach, Irina and Matveev, Rostislav and Spitsyn, Viktor and Makarov, Sergey and Novgorodov, Innokentiy and Osakovsky, Vladimir and Stoneking, Mark and Pakendorf, Brigitte},
  journal={Molecular biology and evolution},
  pages={msw055},
  year={2016},
  publisher={SMBE}
}

%MULTIMIX
@article{churchhouse2013,
  title={Multiway admixture deconvolution using phased or unphased ancestral panels},
  author={Churchhouse, Claire and Marchini, Jonathan},
  journal={Genetic epidemiology},
  volume={37},
  number={1},
  pages={1--12},
  year={2013},
  publisher={Wiley Online Library}

}

@article{eyheramendy2015,
  title={Genetic structure characterization of Chileans reflects historical immigration patterns},
  author={Eyheramendy, Susana and Martinez, Felipe I and Manevy, Federico and Vial, Cecilia and Repetto, Gabriela M},
  journal={Nature communications},
  volume={6},
  year={2015},
  publisher={Nature Publishing Group}
}

%%Five taxon D statistic Dfoil

@article{fontaine2015,
  title={Extensive introgression in a malaria vector species complex revealed by phylogenomics},
  author={Fontaine, Michael C and Pease, James B and Steele, Aaron and Waterhouse, Robert M and Neafsey, Daniel E and Sharakhov, Igor V and Jiang, Xiaofang and Hall, Andrew B and Catteruccia, Flaminia and Kakani, Evdoxia and others},
  journal={Science},
  volume={347},
  number={6217},
  pages={1258524},
  year={2015},
  publisher={American Association for the Advancement of Science}
}

@techreport{terhorst2017,
  title={Robust and scalable inference of population history from hundreds of unphased whole genomes},
  author={Terhorst, Jonathan and Kamm, John A and Song, Yun S},
  year={2017},
  journal={Nature Genetics},
  volume={49},
  pages={303--309}
  institution={Nature Research}
}

@article{pease2016,
  title={Phylogenomics reveals three sources of adaptive variation during a rapid radiation},
  author={Pease, James B and Haak, David C and Hahn, Matthew W and Moyle, Leonie C},
  journal={PLoS Biol},
  volume={14},
  number={2},
  pages={e1002379},
  year={2016},
  publisher={Public Library of Science}
}

%D statistic
@article{durand2011,
  title={Testing for ancient admixture between closely related populations},
  author={Durand, Eric Y and Patterson, Nick and Reich, David and Slatkin, Montgomery},
  journal={Molecular biology and evolution},
  volume={28},
  number={8},
  pages={2239--2252},
  year={2011},
  publisher={SMBE}
}

@article{heliconius2012,
  title={Butterfly genome reveals promiscuous exchange of mimicry adaptations among species},
  author={Heliconius Genome Consortium and others},
  journal={Nature},
  volume={487},
  number={7405},
  pages={94--98},
  year={2012},
  publisher={Nature Publishing Group}
}

%ABC
@article{beaumont2002,
  title={Approximate Bayesian computation in population genetics},
  author={Beaumont, Mark A and Zhang, Wenyang and Balding, David J},
  journal={Genetics},
  volume={162},
  number={4},
  pages={2025--2035},
  year={2002},
  publisher={Genetics Soc America}
}

%dadi
@article{gutenkunst2009,
  title={Inferring the joint demographic history of multiple populations from multidimensional SNP frequency data},
  author={Gutenkunst, Ryan N and Hernandez, Ryan D and Williamson, Scott H and Bustamante, Carlos D},
  journal={PLoS Genet},
  volume={5},
  number={10},
  pages={e1000695},
  year={2009},
  publisher={Public Library of Science}
}

%IM isolation with migration
@article{hey2004,
  title={Multilocus methods for estimating population sizes, migration rates and divergence time, with applications to the divergence of Drosophila pseudoobscura and D. persimilis},
  author={Hey, Jody and Nielsen, Rasmus},
  journal={Genetics},
  volume={167},
  number={2},
  pages={747--760},
  year={2004},
  publisher={Genetics Soc America}
}

%multiple sequentially Markovian coalescent (MSMC)
@article{schiffels2014,
  title={Inferring human population size and separation history from multiple genome sequences},
  author={Schiffels, Stephan and Durbin, Richard},
  journal={Nature genetics},
  volume={46},
  number={8},
  pages={919--925},
  year={2014},
  publisher={Nature Publishing Group}
}

%RASPberry
@article{wegmann2011,
  title={Recombination rates in admixed individuals identified by ancestry-based inference},
  author={Wegmann, Daniel and Kessner, Darren E and Veeramah, Krishna R and Mathias, Rasika A and Nicolae, Dan L and Yanek, Lisa R and Sun, Yan V and Torgerson, Dara G and Rafaels, Nicholas and Mosley, Thomas and others},
  journal={Nature genetics},
  volume={43},
  number={9},
  pages={847--853},
  year={2011},
  publisher={Nature Publishing Group}
}

@article{christe2016,
  title={Selection against recombinant hybrids maintains reproductive isolation in hybridizing Populus species despite F1 fertility and recurrent gene flow},
  author={Christe, Camille and St{\"o}lting, Kai N and Bresadola, Luisa and Fussi, Barbara and Heinze, Berthold and Wegmann, Daniel and Lexer, Christian},
  journal={Molecular ecology},
  volume={25},
  number={11},
  pages={2482--2498},
  year={2016},
  publisher={Wiley Online Library}
}

%LAMP
@article{sankararaman2008,
  title={Estimating local ancestry in admixed populations},
  author={Sankararaman, Sriram and Sridhar, Srinath and Kimmel, Gad and Halperin, Eran},
  journal={The American Journal of Human Genetics},
  volume={82},
  number={2},
  pages={290--303},
  year={2008},
  publisher={Elsevier}
}

@article{patterson2012,
  title={Ancient admixture in human history},
  author={Patterson, Nick and Moorjani, Priya and Luo, Yontao and Mallick, Swapan and Rohland, Nadin and Zhan, Yiping and Genschoreck, Teri and Webster, Teresa and Reich, David},
  journal={Genetics},
  volume={192},
  number={3},
  pages={1065--1093},
  year={2012},
  publisher={Genetics Soc America}
}

%Gmin
@article{kingan2015,
  title={Genome divergence and gene flow between Drosophila simulans and D. mauritiana},
  author={Kingan, Sarah B and Geneva, Anthony J and Vedanayagam, Jeffrey P and Garrigan, Daniel},
  journal={bioRxiv},
  pages={024711},
  year={2015},
  publisher={Cold Spring Harbor Labs Journals}
}

%fineSTRUCTURE
@article{lawson2012,
  title={Inference of population structure using dense haplotype data},
  author={Lawson, Daniel John and Hellenthal, Garrett and Myers, Simon and Falush, Daniel},
  journal={PLoS Genet},
  volume={8},
  number={1},
  pages={e1002453},
  year={2012},
  publisher={Public Library of Science}
}

@article{skoglund2015,
  title={Genetic evidence for two founding populations of the Americas},
  author={Skoglund, Pontus and Mallick, Swapan and Bortolini, Maria C{\'a}tira and Chennagiri, Niru and H{\"u}nemeier, T{\'a}bita and Petzl-Erler, Maria Luiza and Salzano, Francisco Mauro and Patterson, Nick and Reich, David},
  journal={Nature},
  volume={525},
  number={7567},
  pages={104--108},
  year={2015},
  publisher={Nature Publishing Group}
}

%globetrotter
@article{hellenthal2014,
  title={A genetic atlas of human admixture history},
  author={Hellenthal, Garrett and Busby, George BJ and Band, Gavin and Wilson, James F and Capelli, Cristian and Falush, Daniel and Myers, Simon},
  journal={Science},
  volume={343},
  number={6172},
  pages={747--751},
  year={2014},
  publisher={American Association for the Advancement of Science}
}

@article{scott2016,
  title={Characterization of Greater Middle Eastern genetic variation for enhanced disease gene discovery},
  author={Scott, Eric M and Halees, Anason and Itan, Yuval and Spencer, Emily G and He, Yupeng and Azab, Mostafa Abdellateef and Gabriel, Stacey B and Belkadi, Aziz and Boisson, Bertrand and Abel, Laurent and others},
  journal={Nature Genetics},
  year={2016},
  publisher={Nature Publishing Group}
}

@article{pritchard2000,
  title={Inference of population structure using multilocus genotype data},
  author={Pritchard, Jonathan K and Stephens, Matthew and Donnelly, Peter},
  journal={Genetics},
  volume={155},
  number={2},
  pages={945--959},
  year={2000},
  publisher={Genetics Soc America}
}

@article{racimo2016,
  title={Signatures of archaic adaptive introgression in present-day human populations},
  author={Racimo, Fernando and Marnetto, Davide and Huerta-S{\'a}nchez, Emilia},
  journal={Molecular Biology and Evolution},
  pages={msw216},
  year={2016},
  publisher={SMBE}
}

@article{sams2016,
  title={Adaptively introgressed Neandertal haplotype at the OAS locus functionally impacts innate immune responses in humans},
  author={Sams, Aaron J and Dumaine, Anne and N{\'e}d{\'e}lec, Yohann and Yotova, Vania and Alfieri, Carolina and Tanner, Jerome E and Messer, Philipp W and Barreiro, Luis B},
  journal={Genome Biology},
  volume={17},
  number={1},
  pages={246},
  year={2016},
  publisher={BioMed Central}
}

@article{lam2010resequencing,
  title={Resequencing of 31 wild and cultivated soybean genomes identifies patterns of genetic diversity and selection},
  author={Lam, Hon-Ming and Xu, Xun and Liu, Xin and Chen, Wenbin and Yang, Guohua and Wong, Fuk-Ling and Li, Man-Wah and He, Weiming and Qin, Nan and Wang, Bo and others},
  journal={Nature genetics},
  volume={42},
  number={12},
  pages={1053--1059},
  year={2010},
  publisher={Nature Publishing Group}
}

@article{myles2011genetic,
  title={Genetic structure and domestication history of the grape},
  author={Myles, Sean and Boyko, Adam R and Owens, Christopher L and Brown, Patrick J and Grassi, Fabrizio and Aradhya, Mallikarjuna K and Prins, Bernard and Reynolds, Andy and Chia, Jer-Ming and Ware, Doreen and others},
  journal={Proceedings of the National Academy of Sciences},
  volume={108},
  number={9},
  pages={3530--3535},
  year={2011},
  publisher={National Acad Sciences}
}

@article{diez2015olive,
  title={Olive domestication and diversification in the Mediterranean Basin},
  author={Diez, Concepcion M and Trujillo, Isabel and Martinez-Urdiroz, Nieves and Barranco, Diego and Rallo, Luis and Marfil, Pedro and Gaut, Brandon S},
  journal={New Phytologist},
  volume={206},
  number={1},
  pages={436--447},
  year={2015},
  publisher={Wiley Online Library}
}

@article{whitney2006adaptive,
  title={Adaptive introgression of herbivore resistance traits in the weedy sunflower Helianthus annuus},
  author={Whitney, Kenneth D and Randell, Rebecca A and Rieseberg, Loren H},
  journal={The American Naturalist},
  volume={167},
  number={6},
  pages={794--807},
  year={2006},
  publisher={The University of Chicago Press}
}

@article{whitney2010adaptive,
  title={Adaptive introgression of abiotic tolerance traits in the sunflower Helianthus annuus},
  author={Whitney, Kenneth D and Randell, Rebecca A and Rieseberg, Loren H},
  journal={New Phytologist},
  volume={187},
  number={1},
  pages={230--239},
  year={2010},
  publisher={Wiley Online Library}
}

@article{scascitelli2010genome,
  title={Genome scan of hybridizing sunflowers from Texas (Helianthus annuus and H. debilis) reveals asymmetric patterns of introgression and small islands of genomic differentiation},
  author={Scascitelli, M and Whitney, KD and Randell, RA and King, Matthew and Buerkle, CA and Rieseberg, LH},
  journal={Molecular Ecology},
  volume={19},
  number={3},
  pages={521--541},
  year={2010},
  publisher={Wiley Online Library}
}

@article{rieseberg2007hybridization,
  title={Hybridization and the colonization of novel habitats by annual sunflowers},
  author={Rieseberg, Loren H and Kim, Seung-Chul and Randell, Rebecca A and Whitney, Kenneth D and Gross, Briana L and Lexer, Christian and Clay, Keith},
  journal={Genetica},
  volume={129},
  number={2},
  pages={149--165},
  year={2007},
  publisher={Springer}
}

@article{steele2006marker,
  title={Marker-assisted selection to introgress rice QTLs controlling root traits into an Indian upland rice variety},
  author={Steele, KA and Price, AH and Shashidhar, HE and Witcombe, JR},
  journal={Theoretical and Applied Genetics},
  volume={112},
  number={2},
  pages={208--221},
  year={2006},
  publisher={Springer}
}

@article{heiser1951hybridization,
  title={Hybridization in the annual sunflowers: Helianthus annuus$\times$ H. debilis var. cucumerifolius},
  author={Heiser Jr, Charles B},
  journal={Evolution},
  pages={42--51},
  year={1951},
  publisher={JSTOR}
}

@article{dreiseitl2017heterogeneity,
  title={Heterogeneity of Powdery Mildew Resistance Revealed in Accessions of the ICARDA Wild Barley Collection},
  author={Dreiseitl, Antonin},
  journal={Frontiers in Plant Science},
  volume={8},
  year={2017},
  publisher={Frontiers Media SA}
}

@article{bredeson2016sequencing,
  title={Sequencing wild and cultivated cassava and related species reveals extensive interspecific hybridization and genetic diversity},
  author={Bredeson, Jessen V and Lyons, Jessica B and Prochnik, Simon E and Wu, G Albert and Ha, Cindy M and Edsinger-Gonzales, Eric and Grimwood, Jane and Schmutz, Jeremy and Rabbi, Ismail Y and Egesi, Chiedozie and others},
  journal={Nature biotechnology},
  volume={34},
  number={5},
  pages={562--570},
  year={2016},
  publisher={Nature Publishing Group}
}

@article{cornille2012new,
  title={New insight into the history of domesticated apple: secondary contribution of the European wild apple to the genome of cultivated varieties},
  author={Cornille, Amandine and Gladieux, Pierre and Smulders, Marinus JM and Rold{\'a}n-Ruiz, Isabel and Laurens, Fran{\c{c}}ois and Le Cam, Bruno and Nersesyan, Anush and Clavel, Joanne and Olonova, Marina and Feugey, Laurence and others},
  journal={PLoS Genet},
  volume={8},
  number={5},
  pages={e1002703},
  year={2012},
  publisher={Public Library of Science}
}

@article{purugganan2011archaeological,
  title={Archaeological data reveal slow rates of evolution during plant domestication},
  author={Purugganan, Michael D and Fuller, Dorian Q},
  journal={Evolution},
  volume={65},
  number={1},
  pages={171--183},
  year={2011},
  publisher={Wiley Online Library}
}